---
params: 
  chapter: 1
  species: "walleye pollock"
  species_complex: "no"
  area: "Gulf of Alaska"
  tier: 3 
author: "Cole C. Monnahan, Bridget E. Ferriss, S. Kalei Shotwell, Zack Oyafuso, Mike Levine, James T. Thorson, Lauren Rogers, Jane Sullivan, Juliette Champagnat"
date: "November 2024"

# don't do anything below this line unless you are cool with the consequences
# use html_document2 for kniting to html which is better during development
# use word_document2 for final knitting
output: 
  bookdown::word_document2:
     reference_docx: "template.docx"
     toc: false
     number_sections: false
     css: "styles.css"
     pandoc_args:  
       - '--lua-filter=scholarly-metadata.lua'
       - '--lua-filter=author-info-blocks.lua'

always_allow_html: true        
fontsize: 11pt
geometry:
  - top=1in
  - bottom=1in
  - left=1in
  - right=1in     
fig_caption: true
bibliography: "safe_bibby.bib"
csl: "fishfish.csl"
---

```{r setup, include=FALSE, echo = FALSE}
# remotes::install_github("crsh/rmdfiltr")
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "!h", out.extra = "")
knitr::opts_knit$set(eval.after = "fig.cap")
x = 1
sprintf("Figure %s. " , paste0(params$chapter, '.',x))
```

```{r}
if(params$species_complex=="no"){
   title = paste0(params$chapter, ". Assessment of the ", tools::toTitleCase(params$species), " Stock in the ", params$area)
  } else {
    title = paste0(params$chapter, ". Assessment of the ", tools::toTitleCase(params$species), " Stock Complex in the ", params$area)  
  }


```

---
title: `r title`
---

```{r load, warning=F, message=F}
# load
library(safe)
suppressWarnings(library(flextable))
set_flextable_defaults(font.size = 10, padding.top = 0, padding.bottom=0, table.layout='autofit', font.family='Times New Roman')
suppressWarnings(library(readxl))
## library(dplyr)
## library(tidyr)
source("C:/Users/cole.monnahan/Work/assessments/GOA_pollock/2024/writeup/markdown/prepare_global_workspace.R")

```
<p style="text-align:center;">This report may be cited as:</p>

Monnahan, C.C., B. Ferriss, S.K. Shotwell, Z. Oyafuso, M. Levine,
J.T. Thorson, L. Rogers, J. Sullivan and J. Champagnat. 2024.
Assessment of the Walleye Pollock Stock in the Gulf of Alaska.
North Pacific Fishery Management Council, Anchorage, AK.
Available from https://www.npfmc.org/library/safe-reports/

The following documents are associated with this report and can
be found at the associated links:




**Appendix 1.A** is available online [here](https://apps-afsc.fisheries.noaa.gov/Plan_Team/2024/goapollock_appA.pdf).

\pagebreak

```{r, child="01-summary.Rmd", eval = T}
```

```{r, child="02-responses.Rmd", eval = T}
```

```{r, child="03-intro.Rmd", eval = T}
```

```{r, child="04-fishery.Rmd", eval = T}
```

```{r, child="05-data.Rmd", eval = T}
```

```{r, child="06-model.Rmd", eval = T}
```

```{r, child="07-results.Rmd", eval = T}
```

```{r, child="08-abc.Rmd", eval = T}
```

```{r, child="09a-risk-table.Rmd", eval = T}
```

```{r, child="09b-projections.Rmd", eval = T}
```

```{r, child="10-gaps.Rmd", eval = T}
```


```{r, child="11-tables.Rmd", eval = T}
```
 
```{r, child="12-figures.Rmd", eval = T}
```

\pagebreak

<!-- \allsectionsfont{\centering} -->

# References

::: {#refs}
:::

<!-- \pagebreak -->

<!-- ```{r, child="11-tables.Rmd", eval = T} -->

<!-- ```  -->

<!-- \pagebreak -->

<!-- ```{r, child="12-figures.Rmd", eval = T} -->

<!-- ``` -->

\pagebreak

# Appendix 1A. Ecosystem and Socioeconomic Profile of the Walleye Pollock stock in the Gulf of Alaska - Report Card

Appendix 1.A is available at [this external link](https://apps-afsc.fisheries.noaa.gov/Plan_Team/2024/goapollock_appA.pdf).
