# Executive Summary

## Summary of Changes in Assessment Inputs

### Changes to input data
1. Fishery: 2022 total catch was updated and catch at age
   added. The 2023 TAC was used for catch in 2023.
2. Shelikof Strait acoustic survey: 2023 biomass index and age composition.
3. NMFS bottom trawl survey: 2023 index and length compositions
4. Summer acoustic survey: 2023 index and length compositions
5. ADF&G crab/groundfish trawl survey: 2023 biomass index 

### Changes in assessment methodology
This year there was no change in model structure, but there was a
transition to a new modeling platform named Template Model
Builder (TMB; @kristensen2016a). A bridging analysis showed
that the former 19.1a ADMB model and TMB model were equivalent,
with differences of less than 0.02% in estimates and uncertainty
of SSB and recruitment. Given a change in modeling platform, the
previous model 19.1a was renamed to 23.0. Moving to TMB allows
for more flexibility in modeling process errors due to its
ability to efficiently apply the Laplace approximation to get the
marginal likelihood in complex, non-linear hierarchical
models. We explored a suite of alternative fisheries selectivity
models (Appendix 1F) but did not bring them forward for
consideration this year. The advantages of this flexibility will
be explored in depth in future years, but the focus this year was
to transition the assessment to the new platform.

## Summary of Results

The base model projection of female spawning biomass in `r year+1`
is `r ssb` t, which is `r round(100*Finv(ssb)/Finv(b100),1)`% of
unfished spawning biomass (based on average post-1977
recruitment) and above B40% (`r b40` t), thereby placing GOA
pollock in sub-tier "a" of Tier 3. New surveys in 2023 include
the winter Shelikof Strait acoustic survey, summer acoustic
survey, summer NMFS bottom trawl survey, and the ADF&G bottom
trawl survey. These survey indices showed somewhat divergent trends,
with large increases in the summer NMFS bottom trawl (79.4%) and
summer acoustic (71.7%) from 2021, but decreases in the winter
acoustic (-29.2%) and ADF&G bottom trawl survey (-22.5%)
from 2022. Together the new data led to an increased spawning
population relative to the prediction from last year.

The risk matrix table recommended by the Scientific and
Statistical Committee (SSC) was used to determine whether to
recommend an ABC lower than the maximum permissible. The table is
applied by evaluating the severity of four types of
considerations that could be used to support a scientific
recommendation to reduce the ABC from the maximum permissible. We
identified some elevated concerns about the stock assessment,
but none for population dynamics, environment/ecosystem,
or fisheries performance categories. We therefore recommend no
reduction from maximum permissible ABC.

The recommended `r year+1` ABC for pollock in the Gulf of Alaska
west of 140° W lon. (W/C/WYK regions) is `r abc` t, which is an
increase of `r round(100*(abcN-abcN.last)/abcN.last,1)`% from the
`r year` ABC. The recommended `r year+2` ABC is 
`r F(exec_table[12,4])` t. The OFL in `r year+1` is `r ofl` t, and
the OFL in `r year+2` if the ABC is taken in `r year+1` is 
`r F(exec_table[10,4])` t. These calculations are based on a
projected `r year` catch of `r F(tail(datfile$cattot,1))` t and
the ABC for years `r year+1` and `r year+2`. The estimated scale
of the stock increased about 40% compared to previous years,
driven by new data, particularly the new high biomass indices
from summer surveys.

For pollock in southeast Alaska (Southeast Outside region, east
of 140° W lon.), the ABC recommendation for both `r year+1` and
`r year+2` is 9,749 t (see Appendix 1B) and the OFL
recommendation for both `r year+1` and `r year +2` is 12,998
t. These recommendations are based on a Tier 5 assessment using
the projected biomass in 2024 and 2025 from a random effects
model fit to the 1990-2023 bottom trawl survey biomass estimates
of the assessment area. 

\pagebreak
```{r, child= "01a-main_table.Rmd", eval = T}

```



<!-- Text table of area apportionments (if any) for the recommended one- and two-year ahead ABCs and OFLs, with a brief description of the apportionment methodology. -->

```{r, child= "01b-apportionment.Rmd", eval = T}

```
\pagebreak

<!-- If so are so kind as to provide tables for the plan team, place them here -->

<!-- ```{r, child= "01c-plan_team.Rmd", eval = T} -->

<!-- ``` -->


