# Executive Summary

## Summary of Changes in Assessment Inputs

### Changes to input data
 1. Fishery: 2023 total catch was updated and catch at age added.
2024 catch was assumed to be 131,000 t, well below the TAC due a
closure in the Central GOA.
 2. Shelikof Strait acoustic survey: 2024 biomass index and age compositions.
 3. NMFS bottom trawl survey: 2023 age compositions
 4. Summer acoustic survey: 2023 age compositions
 5. ADF&G crab/groundfish trawl survey: 2024 biomass index 
 6. The NMFS bottom trawl age compositions and weight at age were
updated to include only survey tows west of 140W, bringing these
data in line with the biomass index.


### Changes in assessment methodology
Four alternative models were developed in 2024.

1. *Model 23a*: Data weights were revised this year. CVs for
biomass indices and input sample sizes for age compositions for
Shelikof Strait survey, NMFS bottom trawl survey, and summer
acoustic survey were updated.
2. *Model 23b*: Extends model 23a to inncorporate an
environmental covariate on catchability for the Shelikof Strait
survey, based on @rogers2024.
3. *Model 23c*: The same as model 23b but the age 1 and 2 indices
from the Shelikof Strait survey were removed from the model due
to recent poor performance.
4. *Model 23d* **(author recommended)**: Extends model 23c to use
the Dirichlet-multinomial likelihood in place of the multinomial
for all age compositions.

A research model is presented in Appendix 1E, but not proposed as
an alternative this year. This model incorporates a subset of the
ESP data (Appendix 1A) into the assessment to explain variation
in recruitment using a casual framework modeled via dynamic
structural equation models [DSEM, @thorson2024].

## Summary of Results

The following table shows key management quantities for the
alternative models.

```{r abc-table}
nov_abc_table[-5,] %>%
  dplyr::select(version, ssb, B100, fofl, fabc, ofl, abc) %>%
  flextable() %>%
  colformat_double(digits=0) %>%
  colformat_double(digits=3, j=c(4,5)) %>%
  set_header_labels(values=list(version='Model Version', ssb='2025 SSB (t)', B100='B100',
                      fofl='FOFL',fabc='FABC',ofl='OFL', abc='ABC'))
```


The base model (23d) projection of female spawning biomass in
`r year+1` is `r ssb` t, which is 
`r round(100*Finv(ssb)/Finv(b100),1)`% of unfished spawning biomass
(based on average post-1977 recruitment) and above B40% 
(`r b40` t), thereby placing GOA pollock in sub-tier "a" of Tier 3. New
surveys in 2024 include the winter Shelikof Strait acoustic
survey and the ADF&G bottom trawl survey. These survey indices
showed similar trends, with increases in the winter acoustic
(12.0%) and ADF&G bottom trawl survey (17.3%) from 2023. 

The risk matrix table recommended by the Scientific and
Statistical Committee (SSC) was used to determine whether to
recommend an ABC lower than the maximum permissible. The table is
applied by evaluating the severity of four types of
considerations that could be used to support a scientific
recommendation to reduce the ABC from the maximum permissible. We
identified some elevated concerns about the stock assessment,
but none for population dynamics, environment/ecosystem,
or fisheries performance categories. We therefore recommend no
reduction from maximum permissible ABC.

The recommended `r year+1` ABC for pollock in the Gulf of Alaska
west of 140° W lon. (W/C/WYK regions) is `r abc` t, which is an
decrease of `r round(100*(abcN-abcN.last)/abcN.last,1)`% from the
`r year` ABC. The recommended `r year+2` ABC is 
`r F(exec_table[12,4])` t. The OFL in `r year+1` is `r ofl` t, and
the OFL in `r year+2` if the ABC is taken in `r year+1` is 
`r F(exec_table[10,4])` t. These calculations are based on a
projected `r year` catch of `r F(tail(datfile$cattot,1))` t and
the ABC for years `r year+1` and `r year+2`. 

For pollock in southeast Alaska (Southeast Outside region, east
of 140° W lon.), the ABC recommendation for both `r year+1` and
`r year+2` is 9,749 t (see Appendix 1B) and the OFL
recommendation for both `r year+1` and `r year +2` is 12,998
t. These recommendations are based on a Tier 5 assessment using
the projected biomass in 2025 and 2026 from a random effects
model fit to the 1990-2023 bottom trawl survey biomass estimates
of the assessment area. 

\pagebreak
```{r, child= "01a-main_table.Rmd", eval = T}

```



<!-- Text table of area apportionments (if any) for the recommended one- and two-year ahead ABCs and OFLs, with a brief description of the apportionment methodology. -->

```{r, child= "01b-apportionment.Rmd", eval = T}

```


<!-- If so are so kind as to provide tables for the plan team, place them here -->

<!-- ```{r, child= "01c-plan_team.Rmd", eval = T} -->

<!-- ``` -->


