# Data

The data used in the assessment model consist of estimates of
annual catch in tons, fishery age compositions, NMFS summer bottom
trawl survey estimates of biomass and age and length compositions,
acoustic survey estimates of biomass and age composition in
Shelikof Strait, summer acoustic survey estimates of biomass and
age and length composition, and ADF&G bottom trawl survey
estimates of biomass and age composition (Figure
1.\@ref(fig:data-overview)). Binned length composition data are
used in the model only when age composition estimates are
unavailable. The following table specifies the data that were
used in the GOA pollock assessment:

```{r}

data.frame(Source = c(rep("Fishery", 2), 
           rep("Shelikof Strait acoustic survey", 2),
		   rep("Summer acoustic survey", 2),
		   rep("NMFS bottom trawl survey", 2),
  		   rep("ADF&G trawl survey", 2)),
           Data = c("Total catch", "Age composition", "Biomass", "Age composition", 'Biomass', 'Age composition', 'Area-swept biomass', 'Age composition', 'Delta-GLM index', 'Age composition'),
           Years =
    c('1970-2024','1970-2023','1992-2024','1992-2024',
    '2013-2023, biennially',    '2013-2023, biennially',
	'1990-2023, biennially', '1990-2023, biennially',
	'1988-2024', '2000-2022, biennially')) %>%
  flextable::flextable() %>% 
##    flextable::width(j = ~ Source, width = 1) %>% 
##    flextable::width(j = ~ Data, width = 1.5) %>% 
##   flextable::width(j = ~ Years, width = 4) %>% 
##   flextable::merge_v(j = 1) %>% 
  flextable::font(fontname = "Times", part = "all") %>% 
  flextable::fontsize(size = 10) %>% 
  flextable::theme_vanilla() %>% 
  flextable::fix_border_issues() 
```

## Fishery

### Catch

Total catch estimates were obtained from INPFC and ADF&G
publications, and databases maintained at the Alaska Fisheries
Science Center and the Alaska Regional Office. Foreign catches
for 1963-1970 are reported in @forrester1978a. During this period
only Japanese vessels reported catch of pollock in the GOA,
though there may have been some catches by Soviet Union
vessels. Foreign catches 1971-1976 are reported by
@forrester1983a. During this period there are reported pollock
catches for Japanese, Soviet Union, Polish, and South Korean
vessels in the Gulf of Alaska. Foreign and joint venture catches
for 1977-1988 are a blend of estimates from the NORPAC database
maintained by the Alaska Fisheries Science Center. Domestic
catches for 1970-1980 are reported in @rigby1984a. Domestic
catches for 1981-1990 were obtained from PacFIN (Brad Stenberg,
pers. comm. Feb 7, 2014). A discard ratio (discard/retained) of
13.5% was assumed for all domestic catches prior to 1991 based on
the 1991-1992 average discard ratio. Estimated catch for
1991-2020 was obtained from the Catch Accounting System database
maintained by the Alaska Regional Office. These estimates are
derived from shoreside electronic logbooks and observer estimates
of at-sea discards (Table 1.\@ref(tab:retained)). Catches include
the state-managed pollock fishery in Prince William Sound
(PWS). Since 1996, the pollock Guideline Harvest Level (GHL) of
2.5% for the PWS fishery has been deducted from the total
Acceptable Biological Catch (ABC) by the NPFMC Gulf of Alaska
Plan Team for management purposes (see SAFE introduction for
further information). Non-commercial catches are reported in
Table 1.\@ref(tab:non-commercial).

### Age and Size Composition

Catch at age was re-estimated in the 2014 assessment for
1975-1999 from primary databases maintained at AFSC. A simple
non-stratified estimator was used, which consisted of compiling a
single age-length key for use in every year and then applying the
annual length composition to that key. Use of an age-length key
was considered necessary because observers used length-stratified
sampling designs to collect otoliths prior to 1999
[@barbeaux2005a]. Estimates were made separately for the
foreign/JV and domestic fisheries in 1987 when both fisheries
were sampled. There were no major discrepancies between the
re-estimated age composition and estimates that have built up
gradually from assessment to assessment.

Estimates of fishery age composition from 2000 onwards were
derived from at-sea and port sampling of the pollock catch for
length and ageing structures (otoliths). The length composition
and ageing data were obtained from the NORPAC database maintained
at AFSC. Catch age composition was estimated using methods
described by @kimura1989b. Age samples were used to construct
age-length keys by sex and stratum. These keys were applied to
sex and stratum specific length frequency data to estimate age
composition, which were then weighted by the catch in numbers in
each stratum to obtain an overall age composition. A background
age-length key is used fill the gaps in age-length keys by sex
and stratum. Sampling levels by stratum for 2000-2015 are
documented in the assessments available online at
http://www.afsc.noaa.gov/REFM/stocks/Historic_Assess.htm.  Age
and length samples from the `r year-1` fishery were stratified by
half-year seasons and statistical area as follows:


```{r} 
 x <- read_xlsx(path=table.path, sheet = 'fishery_details', range='A1:F7') 
 flextable(x) %>%   
 colformat_double(digits=0, big.mark=',')
``` 

 <!-- ```{r} -->
 <!-- data.frame(age=1:10, Ecatp=round(t(tail(repfile$Ecatp,1)),3)) %>%  -->
 <!--   mutate(cohort=2024-age) -->
 <!-- ``` -->

The dominant cohort in the `r year` expected age composition data
was 2020 with 43%, followed by 2017 and 2018 with 12% and 17%,
respectively. The 2012 cohort is in the plus group and only
accounts for about 4% of expected catch this year.  Fishery catch
at age in 1975-`r year-1` is presented in Table 1.\@ref(tab:caa)
(see also Fig. 1.\@ref(fig:ages-fsh)). Sample sizes for ages and
lengths are given in Table 1.\@ref(tab:aged-fsh).

## Gulf of Alaska Bottom Trawl Survey

Trawl surveys have been conducted by Alaska Fisheries Science
Center (AFSC) beginning in 1984 to assess the abundance of
groundfish in the Gulf of Alaska (Tables 1.\@ref(tab:indices) and
1.\@ref(tab:aged-srv2)). Following standard AFSC practice surveys
prior to 1990 are excluded due to the difficulty in standardizing
the surveys when Japanese vessels with different gear were used.
Starting in 2001, the survey frequency was increased from once
every three years to once every two years. The survey uses a
stratified random design, with 49 strata based on depth, habitat,
and statistical area [@szalay2010a]. Area-swept biomass estimates
are obtained using mean CPUE (standardized for trawling distance
and mean net width) and stratum area. The survey is conducted
from chartered commercial bottom trawlers using standardized
poly-Nor'eastern high opening bottom trawls rigged with roller
gear. In a full three-boat survey, 800 tows are completed, but
the recent average has been closer to 600 tows. On average, 72%
of these tows contain pollock (Table 1.\@ref(tab:aged-srv2)).
Recent years have dropped stations in deeper water which are
unlikely to affect the index due to pollock typically being in
shallower depths with on average 90.9% above 200 m and 99.6%
above 300 m from 1984-2021.

### Biomass Estimates 
The time series of pollock biomass used in the assessment model
is based on the surveyed area in the Gulf of Alaska west of 140°
W long., obtained by adding the biomass estimates for the
Shumagin-610, Chirikof-620, Kodiak-630 statistical areas, and the
western portion of Yakutat-640 statistical area. Biomass
estimates for the west Yakutat area were obtained by splitting
strata and survey CPUE data at 140° W long. and re-estimating
biomass for west Yakutat following @wakabayashi1985 and using the
'gapindex' R package [@oyafuso2024]. In 2001, when the eastern
Gulf of Alaska was not surveyed, a random effects model was used
to interpolate a value for west Yakutat for use in the assessment
model. In 2024 the size composition, age-length key, age
composition and weight at age estimates from 1990-2024 were also
re-estimated to exclude pollock east of 140W, bringing them in
line with the biomass estimates.

The Alaska Fisheries Science Center’s (AFSC) Resource Assessment
and Conservation Engineering (RACE) Division conducted the 18th
comprehensive bottom trawl survey since 1984 during the summer of
2023 (Fig. 1.\@ref(fig:map-BT-cpue)). The 2023 gulfwide biomass
estimate of pollock was 921,886 t, which is an increase of 74.3%
from the 2021 estimate, which itself was a 72.2% increase from
2019, a sharp increase after the low in 2019. The biomass
estimate for the portion of the Gulf of Alaska west of 140º W
long. used in the assessment model is 887,602 t. The coefficient
of variation (CV) of this estimate was 0.13, which is below the
average of 0.197 for the entire time series. 

### Age Composition 

Estimates of numbers at age from the bottom trawl survey are
obtained from random otolith samples and length frequency samples
(Table 1.\@ref(tab:aged-srv2)). Numbers at age are estimated by
statistical area (Shumagin-610, Chirikof-620, Kodiak-630,
Yakutat-640 and Southeastern-650) using a global age-length key
calculated form specimen data west of 140W in each single year,
and CPUE-weighted length frequency data by statistical area. The
2023 ages (Table 1.\@ref(tab:naa-srv2)) were added in 2024.

## Shelikof Strait Acoustic Survey
Winter acoustic surveys to assess the biomass of pre-spawning
aggregations of pollock in Shelikof Strait have been conducted
annually since 1981 (except 1982, 1987, 1999, and 2011). Only
surveys from 1992 and later are used in the stock assessment due
to the higher uncertainty associated with the acoustic estimates
produced with the Biosonics echosounder used prior to 1992.
Additionally, raw survey data are not easily recoverable for the
earlier acoustic surveys, so there is no way to verify (i.e., to
reproduce) the estimates. In 2008, the noise-reduced R/V Oscar
Dyson became the designated survey vessel for acoustic surveys in
the Gulf of Alaska. Survey methods and results for `r year` are
presented in a NMFS processed report [@levine2024].

### Biomass Estimates 
The `r year` biomass estimate for Shelikof Strait in `r year` for
all fish is 289,973 t, which is a 12.0% percent increase from the
`r year-1` estimate (Fig. 1.\@ref(fig:biomass-area-srv1), Table
1.\@ref(tab:indices)). This estimate accounts for escapement of
fish from the midwater survey trawl used to sample echsign (i.e.,
trawl selectivity) via selectivity curves estimated from
recapture nets attached to the survey trawl
[@williams2011;@honkalehto2024]. Winter 2024 pre-spawning pollock
surveys were also conducted in the Shumagin Islands, Chirikof
shelf break, and Pavlov Bay (Fig. 1.\@ref(fig:biomass-area-srv1).
Further information about the surveys conducted in `r year` can
be found in
@levine2024.

<!-- dropped in 2022 b/c only shelikof
<!-- The following table provides results from the 2022 winter -->
<!-- acoustic surveys: -->

<!-- Biomass in Marmot Bay in 2021 increased by 18% compared to 2019, -->
<!-- the last year it was surveyed. Overall, there appears to be a -->
<!-- concentration of spawning activity in Shelikof Strait compared to -->
<!-- other areas in the Gulf of Alaska, but the reduced survey -->
<!-- coverage outside of Shelikof Strait limits the conclusions that -->
<!-- can be drawn. -->

### Age Composition 
Estimates of numbers at age from the Shelikof Strait acoustic
survey (Table 1.\@ref(tab:naa-srv1), Fig. 1.\@ref(fig:ages-srv1))
were obtained using an age-length key compiled from random
otolith samples and applied to weighted length frequency
samples. Sample sizes for ages and lengths are given in Table
1.\@ref(tab:aged-srv1). Estimates of age composition in Shelikof
Strait in `r year` indicate reduced, but persistent dominance of
the 2012 year class, and a mode of age 4-6 fish, indicating new
year classes are starting to comprise the majority of the
spawning and exploitable portion of the population.

Based on recommendations from the 2012 CIE review, 
an approach was developed to model the age-1 and age-2 pollock estimates
separately from the Shelikof Strait acoustic survey biomass and
age composition. These immature fish are not the main target of
the pre-spawning survey, but age-1 and age-2 pollock are highly
variable and occasionally are very abundant in winter acoustic
surveys. By fitting them separately from the 3+ fish it is
possible to utilize an error distribution that better reflects
that variability. Indices are available for both the Shelikof
Strait and Shumagin surveys, but a longer time series of
net-selectivity corrected indices are available for Shelikof
Strait. In addition, model comparisons in the 2018 assessment
indicates that a slightly better fit could be obtained with only
Shelikof Strait indices. The age-1 and age-2 pollock data provide
an invaluable early source of information for recruitment as they
are the only in-year age 1 estimates available for this stock.

However, in 2023 extremely low estimates of some cohorts
were a major concern and discussion point in the 2023 SAFE
[@monnahan2023a]. Explorations revealed that these issues were
caused by unrealistically small observations of age 1 fish from
Shelikof. While the cohorts appear small or very small,
subsequent years of data were not able to counteract the signal
in the age 1 fish. In 2024 another CIE review was done and this
topic was explored in depth, and ended with a recommendation
to either model the data differently or drop them from the model.
So starting in 2024 with model 23c these indices are not fit in
the model, and future research will identify a more appropriate way
to incorporate these observations into the assessment.

### Spawn timing and availability of pollock to the winter Shelikof survey
The Shelikof Strait winter acoustic survey is timed to correspond
to the aggregation of pre-spawning pollock in Shelikof Strait.
However, the timing of spawning has been found to vary from year
to year, which may affect the availability of pollock to the
survey. Variation in spawn timing is not random, but has been
linked to thermal conditions in March and the age structure of
the spawning stock [@rogers2019a]; spawning tends to occur
earlier when temperatures are warmer and when the spawning stock
is older on average. Greater age diversity also results in a more
protracted spawning period, presumably due to the presence of
both early (old) and late (young) spawners, although this has not
been verified in the field. A new approach to account for the
timing of the survey relative to spawning was implemented in 2024
(model 23b) based on @rogers2024. The covariate used is the logit
of proportion mature females greater than 30 cm from Shelikof.
This covariate was slightly outperformed statistically by another
in @rogers2024, but is easier to operationalize because covariate
estimates are always available when a survey occurs. Following
the approach used by the WHAM platform [@stock2021] an expected
covariate ($\hat{X}$) time-series was estimated with an AR(1)
smoother by treating latent states as random effects in TMB and
estimating a variance and correlation for that process.
Observation errors on the covariate were assumed to be 0.02 to
force $\hat{X}$ to be close to the observations. Future work
could explore annually-varying observation errors derived from
data, or estimating it. The expected catchability in year $y$ is
then calculated as

$$q_1(y)=exp(\mu+\beta \hat{X_y} + z_y)$$

where $z_y$ is a random walk component that was previously used
but was disabled in 2024 (i.e., $z_y=0$ for all $y$) per the
recommendation of the SSC.

This covariate-linked model helps explain the misfit
to high survey estimates from 2015-2019, which was a cause of
concern and led to a reduction in the ABC in 2019.

## Summer Acoustic Survey
Six complete acoustic surveys, in 2013, 2015, 2017, 2019, 2021
and 2023, have been conducted by AFSC on the R/V Oscar Dyson in
the Gulf of Alaska during summer
[@jones2014a;@jones2017a;@jones2019a;@jones-prep;@levine2024;@mcgowan-prep].
The area surveyed covers the Gulf of Alaska shelf and upper slope
and associated bays and troughs, from a westward extent of 170°
W, and extends to an eastward extent of 140° W. Prince William
Sound was also surveyed in 2013, 2015, and 2019. The survey
consists of widely-spaced parallel transects along the shelf, and
more closely spaced transects in troughs, bays, and Shelikof
Strait. Mid-water and bottom trawls are used to identify acoustic
targets. The 2023 biomass estimate for summer acoustic survey is
740,417 t, which is a 71.7% percent increase from the 2021
estimate (Table 1.\@ref(tab:indices)). 2023 age compositions were
added in 2024. Analysis of the 2023 survey was not complicated by
the presence of age-0 pollock, which was a problem in previous
summer acoustic surveys because age-0 pollock backscatter cannot
be readily distinguished from age 1+ pollock [@jones2019a].

In 2023 an issue with vessel noise was identified and required a
minor change in the way the data were processed. The processing
methods used in the survey assume that noise is negligible.
However, in 2023 there was concern that this was no longer the
case due to recent changes in vessel noise (sonar self-noise at
38 kHz at survey speed was ~10 dB or ten-fold higher than in
2022). The effects of noise are depth and density dependent and
are difficult to predict. Signal-to-noise thresholding and noise
correction [@derobertis2007a] was used to exclude pollock
backscatter from areas influenced by noise (i.e., all areas with a
signal-to-noise threshold of <6 dB were removed from the
estimate). This revised processing resulted in total pollock
biomass that was 0.19% less than for uncorrected data, confirming
that noise had only a minor impact on the biomass
estimate. Further details can be found in @mcgowan-prep.


## Alaska Department of Fish and Game Crab/Groundfish Trawl Survey
The Alaska Department of Fish and Game (ADF&G) has conducted
bottom trawl surveys of nearshore areas of the Gulf of Alaska
since 1987 (depths from 10-160 m, median of 64 m in 2024;
Fig. 1.\@ref(fig:adfg-map)). Although these surveys are
designed to monitor population trends of Tanner crab and red king
crab, pollock and other fish are also sampled. Standardized
survey methods using a 400-mesh eastern trawl were employed from
1987 to the present. The survey is designed to sample at fixed
stations from mostly nearshore areas from Kodiak Island to Unimak
Pass, and does not cover the entire shelf area
(Fig. 1.\@ref(fig:adfg-map)). The average number of tows
completed during the survey is 352. On average, about 88% of
these tows contained pollock. Details of the ADF&G trawl gear and
sampling procedures are in @spalinger2012a.

The 2024 area-swept biomass estimate for pollock for the ADF&G
crab/groundfish survey was 66,384 t, an increase of 17.2% from
the 2023 biomass estimate (Table 1.\@ref(tab:indices)). The 2024
pollock estimate for this survey is approximately 74% of the
long-term average.

### Biomass Estimates

A delta GLM model was applied to the ADF&G tow by tow data for
1988-`r year` to obtain standardized annual abundance indices.
Data from all years were filtered to exclude missing latitude and
longitudes and missing tows made in lower Shelikof Strait
(between 154.7° W lon. and 156.7° W lon.) were excluded because
these stations were sampled irregularly. The delta GLM model fit
a separate model to the presence-absence observations and to the
positive observations. A fixed effects model was used with the
year, geographic area, and depth as factors. Strata were defined
according to ADF&G district (Kodiak, Chignik, South Peninsula)
and depth (<30 fm, 30-100 fm, >100 fm). Alternative depth strata
were evaluated previously, and model results were found to be
robust to different depth strata assumptions. The same model
structure was used for both the presence-absence observations and
the positive observations. The assumed likelihoods were binomial
for presence-absence observations and gamma for the positive
observations, after evaluation of several alternatives, including
lognormal, gamma, and inverse Gaussian, and which is in line with
recommendations for index standardization [@thorson2021a]. The
model was fit using 'brms' package in R
[@buerkner2017a;@buerkner2018a], which fits Bayesian non-linear
regression models using the modeling framework Stan
[@stanteam2020a]. Posterior samples were generated from six
chains with 3000 iterations (1000 warmup) and thin rate of three.
The analysis had a minimum effective sample size >1000 and
largest Rhat<1.01 and thus passes standard convergence tests.
Comparison of delta-GLM indices the area-swept
estimates indicated similar trends
(Fig. 1.\@ref(fig:adfg-index)). Variances were based on MCMC
sampling from the posterior distribution, and CVs for the annual
index ranged from 0.10 to 0.17. These values likely understate
the uncertainty of the indices with respect to population trends,
since the area covered by the survey is a relatively small
percentage of the GOA shelf area, and so the CVs are scaled up to
have an average of 0.25.

### Age Compositions
Ages were determined by age readers in the AFSC age and growth
unit from samples of pollock otoliths collected during 2000-
`r tail(datfile$srv3_acyrs,1)` ADF&G surveys in even-numbered years
(average sample size = 584; Table 1.\@ref(tab:paa-srv3),
Fig. 1.\@ref(fig:ages-srv3)). Comparison with fishery age
composition shows that older fish (> age-8) are more common in
the ADF&G crab/groundfish survey. This is consistent with the
assessment model, which estimates a domed-shaped selectivity
pattern for the fishery, but an asymptotic selectivity pattern
for the ADF&G survey.

## Data sets considered but not used
### Egg production estimates of spawning biomass
Estimates of spawning biomass in Shelikof Strait based on egg
production methods were produced during 1981-92 (Table
1.\@ref(tab:indices)). A complete description of the estimation
process is given in @picquelle1993a. Egg production estimates were
discontinued in 1992 because the Shelikof Strait acoustic survey
provided similar information. The egg production estimates are
also not used in the assessment model because the surveys are no
longer being conducted, and because the acoustic surveys in
Shelikof Strait show a similar trend over the period when both
were conducted.

### Pre-1984 bottom trawl surveys
Considerable survey work was carried out in the Gulf of Alaska
prior to the start of the NMFS triennial bottom trawl surveys
in 1984. Between 1961 and the mid-1980s, the most common bottom
trawl used for surveying was the 400-mesh eastern trawl. This
trawl (or variants thereof) was used by IPHC for juvenile halibut
surveys in the 1960s, 1970s, and early 1980s, and by NMFS for
groundfish surveys in the 1970s. @szalay2001a
estimated a fishing power correction (FPC) for the ADF&G 400-mesh
eastern trawl of 3.84 (SE = 1.26), indicating that 400-mesh
eastern trawl CPUE for pollock would need to be multiplied by
this factor to be comparable to the NMFS poly-Nor’eastern trawl.

In most cases, earlier surveys in the Gulf of Alaska were not
designed to be comprehensive, with the general strategy being to
cover the Gulf of Alaska west of Cape Spencer over a period of
years, or to survey a large area to obtain an index for group of
groundfish, i.e., flatfish or rockfish. For example,
@ronholt1978a combined surveys for several years to obtain
gulfwide estimates of pollock biomass for 1973-1976. There are
several difficulties with such an approach, including the
possibility of double-counting or missing a portion of the stock
that happened to migrate between surveyed areas. Due to the
difficulty in constructing a consistent time series, the
historical survey estimates are no longer used in the assessment
model.

Multi-year combined survey estimates indicate a large increase in
pollock biomass in the Gulf of Alaska occurred between the early
1960s and the mid 1970s. Increases in pollock biomass between
the 1960s and 1970s were also noted by @alton1987a. In the 1961
survey, pollock were a relatively minor component of the
groundfish community with a mean CPUE of 16 kg/hr
[@ronholt1978a]. Arrowtooth flounder was the most common
groundfish with a mean CPUE of 91 kg/hr. In the 1973-76 surveys,
the CPUE of arrowtooth flounder was similar to the 1961 survey
(83 kg/hr), but pollock CPUE had increased 20-fold to 321 kg/hr,
and was by far the dominant groundfish species in the Gulf of
Alaska. @mueter2002a also found that pollock was low in the
relative abundance in 1960s, became the dominant species in Gulf
of Alaska groundfish community in the 1970s, and subsequently
declined in relative abundance.

Questions concerning the comparability of pollock CPUE data from
historical trawl surveys with later surveys probably can never be
fully resolved. However, because of the large magnitude of the
change in CPUE between the surveys in the 1960s and the early
1970s using similar trawling gear, the conclusion that there was
a large increase in pollock biomass seems robust. Early
speculation about the rise of pollock in the Gulf of Alaska in
the early 1970s implicated the large biomass removals of Pacific
ocean perch, a potential competitor for euphausiid prey
[@somerton1979a; @alton1987a]. Previous work has focused on
role of climate change [@anderson1999a; @bailey2000a]. These
earlier surveys suggest that population biomass in the 1960s,
prior to large-scale commercial exploitation of the stock, may
have been lower than at any time since then.

## Qualitative Trends
To qualitatively assess recent trends in abundance, each survey
time series was standardized by dividing the annual estimate by
the average since 1990. Shelikof Strait acoustic survey estimates
prior to 2008 were rescaled to be comparable to subsequent
surveys conducted by the R/V Oscar Dyson. Although the indices
are not directly comparable due to selectivity differences and
the considerable variability in each survey time series, a fairly
clear downward trend is evident to 2000, followed by a stable,
though variable, trend to 2008, followed by a strong increase to
2013 (Fig. 1.\@ref(fig:fig-indices)). From 2016 to 2019 there was
a strong divergence among the trends, but the relative abundance
came back into reasonable alignment from 2020-2024 with the
exception of the large estimate of the NMFS bottom trawl survey
in 2023.

Indices derived from fisheries catch data were also evaluated for
trends in biological characteristics
(Fig. 1.\@ref(fig:fsh-chars)). The percent of females in the
catch shows some variability and generally is close to 50-50, but
has been low since 2015. Evaluation of sex ratios by season
indicated that this decrease was mostly due a low percentage of
females during the A season prior to
spawning. However the sex ratio during the B
season was close to 50-50, suggesting the skewed sex in
winter was related to spawning behavior, rather than an
indication of a population characteristic. The mean age shows
interannual variability due to strong year classes passing
through the population, but there are no downward trends that
would suggest excessive mortality rates. The percent of old fish
in the catch (nominally defined as age 8 and older) is also
highly variable due to variability in year class strength. The
percent of old fish declined in 2015-2018 as the strong 2012 year
class recruited to the fishery, but increased when the 2012 year
class became age 8 in 2020. With large incoming cohorts and the
decline of the 2012 cohort, the mean age has begun to decrease
again. Under a constant F40% harvest rate, the mean percent of
age 8 and older fish in the catch would be approximately 8%.

An annual index of catch at age diversity was computed using the
Shannon-Wiener information index, $H'$, defined as

$$H'= -\sum_a{ p_a \ln {p_a}}
$$

where $p_a$ is the proportion at age and higher $H'$ values correspond
to higher diversity. Increases in fishing mortality would tend to
reduce age diversity, but year class variability would also
influence it. Age diversity was relatively stable during
1975-2015, but declined sharply to a low in 2016 and has been
increasing since due to the dominance of the 2012 year class in
the catch (Fig. 1.\@ref(fig:fsh-chars)). In 2021 the age
diversity returned to near the long-term average and remains
there through 2024.

The 2012 year class, which is both very strong, and which has
experienced anomalous environmental conditions during the marine
heatwave in the North Pacific during 2015-2017, has displayed
unusual life history characteristics. These include early
maturation, reduced growth, but apparently not reduced total
mortality. It is unclear whether these changes are a result of
density dependence or environmental forcing. Previous assessments
examined this cohort in more depth, but its impact on the fishery
is diminished and is not a focus here.
