# Data

The data used in the assessment model consist of estimates of
annual catch in tons, fishery age composition, NMFS summer bottom
trawl survey estimates of biomass and age and length composition,
acoustic survey estimates of biomass and age composition in
Shelikof Strait, summer acoustic survey estimates of biomass and
age and length composition, and ADF&G bottom trawl survey
estimates of biomass and age composition. Binned length
composition data are used in the model only when age composition
estimates are unavailable. The following table specifies the data
that were used in the GOA pollock assessment:

```{r}

data.frame(Source = c(rep("Fishery", 2), 
           rep("Shelikof Strait acoustic survey", 2),
		   rep("Summer acoustic survey", 2),
		   rep("NMFS bottom trawl survey", 2),
  		   rep("ADF&G trawl survey", 2)),
           Data = c("Total catch", "Age composition", "Biomass", "Age composition", 'Biomass', 'Age composition', 'Area-swept biomass', 'Age composition', 'Delta-GLM index', 'Age composition'),
           Years =
    c('1970-2021','1970-2021','1992-2022','1992-2022',
    '2013-2021, biennially',    '2013-2021, biennially',
	'1990-2021, biennially', '1990-2021, biennially',
	'1988-2022', '2000-2020, biennially')) %>%
  flextable::flextable() %>% 
##    flextable::width(j = ~ Source, width = 1) %>% 
##    flextable::width(j = ~ Data, width = 1.5) %>% 
##   flextable::width(j = ~ Years, width = 4) %>% 
##   flextable::merge_v(j = 1) %>% 
  flextable::font(fontname = "Times", part = "all") %>% 
  flextable::fontsize(size = 10) %>% 
  flextable::theme_vanilla() %>% 
  flextable::fix_border_issues() 

```

## Fishery

### Catch

Total catch estimates were obtained from INPFC and ADF&G
publications, and databases maintained at the Alaska Fisheries
Science Center and the Alaska Regional Office. Foreign catches
for 1963-1970 are reported in @forrester1978a. During this period
only Japanese vessels reported catch of pollock in the GOA,
though there may have been some catches by Soviet Union
vessels. Foreign catches 1971-1976 are reported by
@forrester1983a. During this period there are reported pollock
catches for Japanese, Soviet Union, Polish, and South Korean
vessels in the Gulf of Alaska. Foreign and joint venture catches
for 1977-1988 are blend estimates from the NORPAC database
maintained by the Alaska Fisheries Science Center. Domestic
catches for 1970-1980 are reported in @rigby1984a. Domestic
catches for 1981-1990 were obtained from PacFIN (Brad Stenberg,
pers. comm. Feb 7, 2014). A discard ratio (discard/retained) of
13.5% was assumed for all domestic catches prior to 1991 based on
the 1991-1992 average discard ratio. Estimated catch for
1991-2020 was obtained from the Catch Accounting System database
maintained by the Alaska Regional Office. These estimates are
derived from shoreside electronic logbooks and observer estimates
of at-sea discards (Table 1.\@ref(tab:retained)). Catches include
the state-managed pollock fishery in Prince William Sound
(PWS). Since 1996, the pollock Guideline Harvest Level (GHL) of
2.5% for the PWS fishery has been deducted from the total
Acceptable Biological Catch (ABC) by the NPFMC Gulf of Alaska
Plan Team for management purposes (see SAFE introduction for
further information). Non-commercial catches are reported in
Appendix 1E.

### Age and Size Composition

Catch at age was re-estimated in the 2014 assessment for
1975-1999 from primary databases maintained at AFSC. A simple
non-stratified estimator was used, which consisted of compiling a
single age-length key for use in every year and the applying the
annual length composition to that key. Use of an age-length key
was considered necessary because observers used length-stratified
sampling designs to collect otoliths prior to 1999
[@barbeaux2005a]. Estimates were made separately for the
foreign/JV and domestic fisheries in 1987 when both fisheries
were sampled. There were no major discrepancies between the
re-estimated age composition and estimates that have built up
gradually from assessment to assessment.

Estimates of fishery age composition from 2000 onwards were
derived from at-sea and port sampling of the pollock catch for
length and ageing structures (otoliths). The length composition
and ageing data were obtained from the NORPAC database maintained
at AFSC. Catch age composition was estimated using methods
described by @kimura1989b. Age samples were used to construct
age-length keys by sex and stratum. These keys were applied to
sex and stratum specific length frequency data to estimate age
composition, which were then weighted by the catch in numbers in
each stratum to obtain an overall age composition. A background
age-length key is used fill the gaps in age-length keys by sex
and stratum. Sampling levels by stratum for 2000-2015 are
documented in the assessments available online at
http://www.afsc.noaa.gov/REFM/stocks/Historic_Assess.htm.  Age
and length samples from the `r year-1` fishery were stratified by
half-year seasons and statistical area as follows:


```{r} 
 x <- read_xlsx(path=table.path, sheet = 'fishery_details', range='A1:F7') 
 flextable(x) %>%   
 colformat_double(digits=0, big.mark=',')  
``` 

The estimated age composition in 2021 in all areas and all
seasons was notable because it was not dominated by age-9 fish
(2012 year class) for the first time in many years (except last
year), particularly for the B season
(Fig. 1.\@ref(fig:catches-map)). Instead, the age-4 fish had the
largest percentage with 38%, age-3 (2018 cohort) with 23%, and
the age-9 fish only accounting for 17%. Younger fish are likely
to become increasingly prominent in the catch-at-age as the 2012
year class begins age out of the population. Fishery catch at age
in 1975-`r year` is presented in Table 1.\@ref(tab:caa) (See also
Fig. 1.\@ref(fig:ages-fsh)). Sample sizes for ages
and lengths are given in Table 1.\@ref(tab:aged-fsh).

## Gulf of Alaska Bottom Trawl Survey

Trawl surveys have been conducted by Alaska Fisheries Science
Center (AFSC) beginning in 1984 to assess the abundance of
groundfish in the Gulf of Alaska (Tables 1.\@ref(tab:indices) and
1.\@ref(tab:aged-srv2)). Starting in 2001, the survey frequency
was increased from once every three years to once every two
years. The survey uses a stratified random design, with 49 strata
based on depth, habitat, and statistical area
[@szalay2010a]. Area-swept biomass estimates are obtained using
mean CPUE (standardized for trawling distance and mean net width)
and stratum area. The survey is conducted from chartered
commercial bottom trawlers using standardized poly-Nor'eastern
high opening bottom trawls rigged with roller gear. In a full
three-boat survey, 800 tows are completed, but the recent average
has been closer to 600 tows. On average, 72% of these tows
contain pollock (Table 1.\@ref(tab:aged-srv2)). Recent years have
dropped stations in deeper water which are unlikely to affect the
index due to pollock typically being in shallower depths with on
average 90.9% below 200 m and 99.6% below 300 m from 1984-2021.

### Biomass Estimates 
The time series of pollock biomass used in the assessment model
is based on the surveyed area in the Gulf of Alaska west of 140°
W long., obtained by adding the biomass estimates for the
Shumagin-610, Chirikof-620, Kodiak-630 statistical areas, and the
western portion of Yakutat-640 statistical area. Biomass
estimates for the west Yakutat area were obtained by splitting
strata and survey CPUE data at 140° W long. and re-estimating
biomass for west Yakutat. In 2001, when eastern Gulf of Alaska
was not surveyed, a random effects model was used to interpolate
a value for west Yakutat for use in the assessment model.

The Alaska Fisheries Science Center’s (AFSC) Resource Assessment
and Conservation Engineering (RACE) Division conducted the
seventeenth comprehensive bottom trawl survey since 1984 during
the summer of 2021 (Fig. 1.\@ref(fig:map-BT-cpue). The 2021
gulfwide biomass estimate of pollock was 528,841 t, which is an
increase of 72.2% from the 2019 estimate, which was the second
lowest in the time series after 2001. The biomass estimate for
the portion of the Gulf of Alaska west of 140º W long. used in
the assessment model is 494,743 t. The coefficient of variation
(CV) of this estimate was 0.17, which is slightly below the
average for the entire time series. Surveys from 1990 onwards are
used in the assessment due to the difficulty in standardizing the
surveys in 1984 and 1987, when Japanese vessels with different
gear were used.

### Age Composition 

Estimates of numbers at age from the bottom trawl survey are
obtained from random otolith samples and length frequency samples
(Table 1.\@ref(tab:aged-srv2)). Numbers at age are estimated by
statistical area (Shumagin-610, Chirikof-620, Kodiak-630,
Yakutat-640 and Southeastern-650) using a global age-length key
for all strata in each single year, and CPUE-weighted length
frequency data by statistical area. The new 2021 combined
Shumagin, Chirikof and Kodiak age composition is used in the
assessment model (Fig. 1.\@ref(fig:ages-srv2), Table
1.\@ref(tab:naa-srv2)), and the 2021 length composotions were
removed. Age-1 pollock (2020 cohort) were strongly present in the
Chirikof, Kodiak, and Yakutat statistical areas, but much less
abundant in the Shumagin and Southeast Alaska areas
(Fig. 1.\@ref(fig:ages-area-srv2)).

## Shelikof Strait Acoustic Survey
Winter acoustic surveys to assess the biomass of pre-spawning
aggregations of pollock in Shelikof Strait have been conducted
annually since 1981 (except 1982, 1987, 1999, and 2011). Only surveys
from 1992 and later are used in the stock assessment due to the
higher uncertainty associated with the acoustic estimates
produced with the Biosonics echosounder used prior
to 1992. Additionally, raw survey data are not easily recoverable
for the earlier acoustic surveys, so there is no way to verify
(i.e., to reproduce) the estimates. Survey methods and results
for 2021 are presented in a NMFS processed report
[@mcgowan-prep]. In 2008, the noise-reduced R/V Oscar Dyson
became the designated survey vessel for acoustic surveys in the
Gulf of Alaska. In winter of 2007, a vessel comparison experiment
was conducted between the R/V Miller Freeman (MF) and the R/V
Oscar Dyson (OD), which obtained an OD/MF ratio of 1.132 for the
acoustic backscatter detected by the two vessels in Shelikof
Strait.

### Biomass Estimates 
The `r year` biomass estimate for Shelikof Strait in `r year` for
all fish is 365,411 t, which is a 30.7% percent decrease from the
2021 estimate (Fig. 1.\@ref(fig:biomass-area-srv1)). This
estimate accounts for trawl selectivity by scaling up the number
of retained pollock by selectivity curves estimated with pocket
nets attached to the midwater trawl used to sample echosign,
continuing an approach that was started in the 2018
assessment. Originally, winter 2022 pre-spawning pollock surveys
were also planned in the Shumagin Islands area, Chirikof shelf
break, Marmot Bay, and Morzhovoi Bay. Due to travel, vessel, and
staffing constraints stemming from protocols required to mitigate
the COVID-19 pandemic, only Shelikof was completed in 2022.

<!-- dropped in 2022 b/c only shelikof
<!-- The following table provides results from the 2022 winter -->
<!-- acoustic surveys: -->

<!-- Biomass in Marmot Bay in 2021 increased by 18% compared to 2019, -->
<!-- the last year it was surveyed. Overall, there appears to be a -->
<!-- concentration of spawning activity in Shelikof Strait compared to -->
<!-- other areas in the Gulf of Alaska, but the reduced survey -->
<!-- coverage outside of Shelikof Strait limits the conclusions that -->
<!-- can be drawn. -->

### Age Composition 
Estimates of numbers at age from the Shelikof Strait acoustic
survey (Table 1.\@ref(tab:naa-srv1), Fig. 1.\@ref(fig:ages-srv1))
were obtained using an age-length key compiled from random
otolith samples and applied to weighted length frequency
samples. Sample sizes for ages and lengths are given Table
1.\@ref(tab:aged-srv1). Estimates of age composition in Shelikof
Strait in `r year` indicate reduced dominance of the 2012 year
class, and a mode of age 2 and 4-5 fish (2020, 2018 and 2017 year
classes), indicating new year classes are starting to comprise
the majority of the spawning and exploitable portion of the
population.

Based on recommendations from the 2012 CIE review, we developed
an approach to model the age-1 and age-2 pollock estimates
separately from the Shelikof Strait acoustic survey biomass and
age composition. These immature fish are not the main target of
the pre-spawning survey, but age-1 and age-2 pollock are highly
variable and occasionally are very abundant in winter acoustic
surveys. By fitting them separately from the 3+ fish it is
possible utilize an error distribution that better reflects that
variability. Indices are available for both the Shelikof Strait
and Shumagin surveys, but a longer time series of net-selectivity
corrected indices are available for Shelikof Strait. In addition,
model comparisons in the 2018 assessment indicates that a
slightly better fit could be obtained with only Shelikof Strait
indices. Therefore this time series was used in the model, but
this decision should be revisited as additional data become
available. The age-2 index in 2020 showed a marked reduction in
comparison to the age-1 index in 2019, which indicated high
abundance of the 2018 year class. Typically year classes that are
abundant in Shelikof Strait at age 1 are also abundant at age 2
in the survey in the following year. The 2018 cohort comprised
15% of the age composition in 2021 (excluding age 1 and 2 fish),
but 29% as 4 year olds in 2022, giving contradictory evidence for
marked decrease from initial estimates as age 1
fish. Consequently, there is considerable uncertainty regarding
the fate of 2018 year class, which may have exited Shelikof
Strait for some reason and be distributed elsewhere in the GOA,
or suffered extremely high mortality. This point is addressed
further in the risk table below.

### Spawn timing and availability of pollock to the winter Shelikof survey
The Shelikof Strait winter acoustic survey is timed to correspond
to the aggregation of pre-spawning pollock in Shelikof
Strait. However, the timing of spawning has been found to vary
from year to year, which may affect the availability of pollock
to the survey. Variation in spawn timing is not random, but has
been linked to thermal conditions in March and the age structure
of the spawning stock [@rogers2019a]; spawning tends to occur
earlier when temperatures are warmer and when the spawning stock
is older on average. Greater age diversity also results in a more
protracted spawning period, presumably due to both early (old)
and late (young) spawners, although this has not been verified in
the field. A new approach was developed in 2022 and shows great
promise, but was not put forward this year for
consideration. Summaries of the work were presented to the Plan
Team and furhter details can be found under the "Models under
development" section of [this document](https://meetings.npfmc.org/CommentReview/DownloadFile?p=7df129bb-16cb-4dba-a4d5-a160c31d087e.pdf&fileName=GOA%20pollock%20model%20updates.pdf).


## Summer Acoustic Survey
Five complete acoustic surveys, in 2013, 2015, 2017, 2019 and
2021, have been conducted by AFSC on the R/V Oscar Dyson in the
Gulf of Alaska during summer
[@jones2014a;@jones2017a;@jones2019a;@jones-prep;@levine-prep]. The
area surveyed covers the Gulf of Alaska shelf and upper slope and
associated bays and troughs, from a westward extent of 170° W
Lon, and extends to an eastward extent of 140° W lon. Prince
William Sound was also surveyed in 2013, 2015, and 2019. The
survey consists of widely-spaced parallel transects along the
shelf, and more closely spaced transects in troughs, bays, and
Shelikof Strait. Mid-water and bottom trawls are used to identify
acoustic targets. The 2021 biomass estimate for summer acoustic
survey is 431,148 t, which is a 25% percent decrease from the
2019 estimate (Table 1.\@ref(tab:indices)). Age composition data
were available in 2022 and showed strong 2017, 2018, and 2020
year classes.  Analysis of the 2019
and 2021 survey was not complicated by the presence of age-0
pollock, which was a problem in previous summer acoustic surveys
because age-0 pollock backscatter cannot be readily distinguished
from age 1+ pollock [@jones2019a].

## Alaska Department of Fish and Game Crab/Groundfish Trawl Survey
The Alaska Department of Fish and Game (ADF&G) has conducted
bottom trawl surveys of nearshore areas of the Gulf of Alaska
since 1987 (depths from 9-137 m, median of 60 m in 2022;
Fig. 1.\@ref(fig:adfg-locations)). Although these surveys are
designed to monitor population trends of Tanner crab and red king
crab, pollock and other fish are also sampled. Standardized
survey methods using a 400-mesh eastern trawl were employed from
1987 to the present. The survey is designed to sample at fixed
stations from mostly nearshore areas from Kodiak Island to Unimak
Pass, and does not cover the entire shelf area
(Fig. 1.\@ref(fig:adfg-locations)). The average number of tows
completed during the survey is 353. On average, about 87% of
these tows contained pollock. Details of the ADF&G trawl gear and
sampling procedures are in Spalinger (2012).

The 2022 area-swept biomass estimate for pollock for the ADF&G
crab/groundfish survey was 71,196 t, and increase of 9.2% from
the 2021 biomass estimate (Table 1.\@ref(tab:indices)). The 2022
pollock estimate for this survey is approximately 78% of the
long-term average.

### Biomass Estimates

A simple delta GLM model was applied to the ADF&G tow by tow data
for 1988-`r year` to obtain annual abundance indices. Data from
all years were filtered to exclude missing latitude and
longitudes and missing tows made in lower Shelikof Strait
(between 154.7° W lon. and 156.7° W lon.) were excluded because
these stations were sampled irregularly. The delta GLM model fit
a separate model to the presence-absence observations and to the
positive observations. A fixed effects model was used with the
year, geographic area, and depth as factors. Strata were defined
according to ADF&G district (Kodiak, Chignik, South Peninsula)
and depth (<30 fm, 30-100 fm, >100 fm). Alternative depth strata
were evaluated previously, and model results were found to be
robust to different depth strata assumptions. The same model
structure was used for both the presence-absence observations and
the positive observations. The assumed likelihoods were binomial
for presence-absence observations and gamma for the positive
observations, after evaluation of several alternatives, including
lognormal, gamma, and inverse Gaussian, and which is in line with
recommendations for index standardization [@thorson2021a]. The
model was fit using 'brms' package in R
[@buerkner2017a;@buerkner2018a], which fits Bayesian non-linear
regression models using the modeling framework Stan
[@stanteam2020a]. Comparison of delta-GLM indices the area-swept
estimates indicated similar trends
(Fig. 1.\@ref(fig:adfg-index)). Variances were based on MCMC
sampling from the posterior distribution, and CVs for the annual
index ranged from 0.10 to 0.17. These values likely understate
the uncertainty of the indices with respect to population trends,
since the area covered by the survey is a relatively small
percentage of the GOA shelf area, and so the CVs are scaled up to
have an average of 0.25.

### Age Compositions
Ages were determined by age readers in the AFSC age and growth
unit from samples of pollock otoliths collected during 2000-
`r tail(datfile$srv3_acyrs,1)` ADF&G surveys in even-numbered years
(average sample size = 583; Table 1.\@ref(tab:paa-srv3),
Fig. 1.\@ref(fig:ages-srv3)). Comparison with fishery age
composition shows that older fish (> age-8) are more common in
the ADF&G crab/groundfish survey. This is consistent with the
assessment model, which estimates a domed-shaped selectivity
pattern for the fishery, but an asymptotic selectivity pattern
for the ADF&G survey.

## Data sets considered but not used
### Egg production estimates of spawning biomass
Estimates of spawning biomass in Shelikof Strait based on egg
production methods were produced during 1981-92 (Table
1.\@ref(tab:indices)). A complete description of the estimation
process is given in @picquelle1993a. Egg production estimates were
discontinued in 1992 because the Shelikof Strait acoustic survey
provided similar information. The egg production estimates are
also not used in the assessment model because the surveys are no
longer being conducted, and because the acoustic surveys in
Shelikof Strait show a similar trend over the period when both
were conducted.

### Pre-1984 bottom trawl surveys
Considerable survey work was carried out in the Gulf of Alaska
prior to the start of the NMFS triennial bottom trawl surveys
in 1984. Between 1961 and the mid-1980s, the most common bottom
trawl used for surveying was the 400-mesh eastern trawl. This
trawl (or variants thereof) was used by IPHC for juvenile halibut
surveys in the 1960s, 1970s, and early 1980s, and by NMFS for
groundfish surveys in the 1970s. @szalay2001a
estimated a fishing power correction (FPC) for the ADF&G 400-mesh
eastern trawl of 3.84 (SE = 1.26), indicating that 400-mesh
eastern trawl CPUE for pollock would need to be multiplied by
this factor to be comparable to the NMFS poly-Nor’eastern trawl.

In most cases, earlier surveys in the Gulf of Alaska were not
designed to be comprehensive, with the general strategy being to
cover the Gulf of Alaska west of Cape Spencer over a period of
years, or to survey a large area to obtain an index for group of
groundfish, i.e., flatfish or rockfish. For example,
@ronholt1978a combined surveys for several years to obtain
gulfwide estimates of pollock biomass for 1973-1976. There are
several difficulties with such an approach, including the
possibility of double-counting or missing a portion of the stock
that happened to migrate between surveyed areas. Due to the
difficulty in constructing a consistent time series, the
historical survey estimates are no longer used in the assessment
model.

Multi-year combined survey estimates indicate a large increase in
pollock biomass in the Gulf of Alaska occurred between the early
1960s and the mid 1970s. Increases in pollock biomass between
the1960s and 1970s were also noted by @alton1987a. In the 1961
survey, pollock were a relatively minor component of the
groundfish community with a mean CPUE of 16 kg/hr
[@ronholt1978a]. Arrowtooth flounder was the most common
groundfish with a mean CPUE of 91 kg/hr. In the 1973-76 surveys,
the CPUE of arrowtooth flounder was similar to the 1961 survey
(83 kg/hr), but pollock CPUE had increased 20-fold to 321 kg/hr,
and was by far the dominant groundfish species in the Gulf of
Alaska. @mueter2002a also found that pollock was low in the
relative abundance in 1960s, became the dominant species in Gulf
of Alaska groundfish community in the 1970s, and subsequently
declined in relative abundance.

Questions concerning the comparability of pollock CPUE data from
historical trawl surveys with later surveys probably can never be
fully resolved. However, because of the large magnitude of the
change in CPUE between the surveys in the 1960s and the early
1970s using similar trawling gear, the conclusion that there was
a large increase in pollock biomass seems robust. Early
speculation about the rise of pollock in the Gulf of Alaska in
the early 1970s implicated the large biomass removals of Pacific
ocean perch, a potential competitor for euphausid prey
[@somerton1979a; @alton1987a]. More recent work has focused on
role of climate change [@anderson1999a; @bailey2000a]. These
earlier surveys suggest that population biomass in the 1960s,
prior to large-scale commercial exploitation of the stock, may
have been lower than at any time since then.

## Qualitative Trends
To qualitatively assess recent trends in abundance, each survey
time series was standardized by dividing the annual estimate by
the average since 1990. Shelikof Strait acoustic survey estimates
prior to 2008 were rescaled to be comparable to subsequent
surveys conducted by the R/V Oscar Dyson. Although there is
considerable variability in each survey time series, a fairly
clear downward trend is evident to 2000, followed by a stable,
though variable, trend to 2008, followed by a strong increase to
2013 (Fig. 1.\@ref(fig:fig-indices)). From 2016 to 2019 there was
a strong divergence among the trends, but with the large
reduction in biomass in 2020 for the Shelikof Strait survey, and
an increase in the ADF&G index, relative abundance has come back
into reasonable alignment since 2020. Indices derived from
fisheries catch data were also evaluated for trends in biological
characteristics (Fig. 1.\@ref(fig:fsh-chars)). The percent of
females in the catch shows some variability but no obvious trend,
and is usually close to 50-50. In 2016, percent female dropped to
40%, but has slowly increased up to 47% in 2022. Evaluation of
sex ratios by season indicated that this decrease was mostly due
a low percentage of females during the A and B seasons prior to
spawning. However the sex ratio during the C and D seasons was
close to 50-50, suggesting the skewed sex in winter was related
to spawning behavior, rather than an indication of a population
characteristic. The mean age shows interannual variability due to
strong year classes passing through the population, but there are
no downward trends that would suggest excessive mortality
rates. The percent of old fish in the catch (nominally defined as
age 8 and older) is also highly variable due to variability in
year class strength. The percent of old fish declined in
2015-2018 as the strong 2012 year class recruited to the fishery,
but increased when the 2012 year class became age 8 in 2020. With
large incoming cohorts and the decline of the 2012 cohort, the
mean age has begun to decrease again. Under a constant F40%
harvest rate, the mean percent of age 8 and older fish in the
catch would be approximately 8%.

An annual index of catch at age diversity was computed using the
Shannon-Wiener information index, $H'$, defined as

$$H'= -\sum_a{ p_a \ln {p_a}}
$$

where $p_a$ is the proportion at age and higher values correspond
to higher diversity. Increases in fishing mortality would tend to
reduce age diversity, but year class variability would also
influence it. Age diversity was relatively stable during
1975-2015, but declined sharply to a low in 2016 and has been
increasing since due to the dominance of the 2012 year class in
the catch (Fig. 1.\@ref(fig:fsh-chars)). In 2021 the age
diversity returned to near the long-term average.

The 2012 year class, which is both very strong, and which has
experienced anomalous environmental conditions during the marine
heatwave in the North Pacific during 2015-2017, has displayed
unusual life history characteristics. These include early
maturation, reduced growth, but apparently not reduced total
mortality (Fig. 1.\@ref(fig:2012-bio)). It is unclear whether
these changes are a result of density dependence or environmental
forcing.
