# Analytical approach

## General Model Structure


An age-structured model covering the period from 1970 to `r year`
(`r length(repfile$years)` years) was used to assess Gulf of Alaska
pollock. The modeled population includes individuals from age 1
to age 10, with age 10 defined as a "plus" group, i.e., all
individuals age 10 and older. Population dynamics were modeled
using standard formulations for mortality and fishery catch
[e.g., @fournier1982a; @deriso1985a; @hilborn1992a]. Year- and
age-specific fishing mortality was modeled as a product of a year
effect, representing the full-selection fishing mortality, and an
age effect, representing the selectivity of that age group to the
fishery. The age effect was modeled using a double-logistic
function with time-varying parameters [@dorn1990a;
@sullivan1997a]. The model was fit to time series of catch
biomass, survey indices of abundance, and estimates of age and
length composition from the fishery and surveys. Details of the
population dynamics and estimation equations are presented in
Appendix 1C.

Model parameters were estimated by maximizing the joint log
likelihood of the data and penalties, viewed as a function of
the parameters. Mean-unbiased log-normal likelihoods were used
for survey biomass and total catch estimates, and multinomial
likelihoods were used for age and length composition data. Model
tuning for composition data was done by iterative re-weighting of
input sample sizes using the @francis2011a method. Variance
estimates/assumptions for survey indices were not reweighted. The
following table lists the likelihood components used in fitting
the model.

```{r}
x <- read_xlsx(table.path, sheet='likelihoods')
flextable(x)
```

### Recruitment
Age composition in the first year is estimated with a single log
deviation for recruitment abundance, which was then decremented
by natural mortality to fill out the initial age vector. In
previous versions of the model, a recruitment penalty of
($\sigma_R =1.0$) was added only to recruitments for 1970-77, and
in the last two years of the model and the rest were were
estimated as free parameters. Starting in 2022 with model 19.1a
the penalty was applied to all deviations, with a value of
$\sigma_R =1.3$ coming from an estimate of a state-space research
version of the model. This change had a relatively small impact on
the estimated recruits and management reference points.

### Modeling fishery data
To accommodate changes in selectivity, we estimated year-specific
parameters for the slope and the intercept parameter for the
ascending logistic portion of selectivity curve (i.e., younger
fish). Variation in these parameters was constrained using a
random walk penalty.

### Modeling survey data
Survey abundance was assumed to be proportional to total
abundance as modified by the estimated survey selectivity
pattern. Expected population numbers at age for the survey were
based on the mid-date of the survey, assuming constant fishing
and natural mortality throughout the year. Standard deviations in
the log-normal likelihood were set equal to the sampling error CV
(coefficient of variation) associated with each survey estimate
of abundance [@kimura1991a].

Survey catchability coefficients can be fixed or freely
estimated. The base model estimated the NMFS bottom trawl survey
catchability, but used a log normal prior with a median of 0.85
and log standard deviation 0.1 based on expert judgment as a
constraint on potential values
(Fig. 1.\@ref(fig:BT-prior)). Catchability coefficients for other
surveys were estimated as free parameters. The age-1 and age-2
winter acoustic survey indices are numerical abundance estimates,
and were modeled using independently estimated catchability
coefficients (i.e., no selectivity is estimated).

A vessel comparison (VC) experiment was conducted in March 2007
during the Shelikof Strait acoustic survey. The VC experiment
involved the R/V Miller Freeman (MF, the survey vessel used to
conduct Shelikof Strait surveys since the mid-1980s), and the R/V
Oscar Dyson (OD), a noise-reduced survey vessel designed to
conduct surveys that have traditionally been done with the R/V
Miller Freeman. The vessel comparison experiment was designed to
collect data either with the two vessels running beside one
another at a distance of 0.7 nmi, or with one vessel following
nearly directly behind the other at a distance of about 1
nmi. The methods were similar to those used during the 2006
Bering Sea VC experiment [@robertis2008a]. Results indicate that
the ratio of 38 kHz pollock backscatter from the R/V Oscar Dyson
relative to the R/V Miller Freeman was significantly greater than
one (1.13), as would be expected if the quieter OD reduced the
avoidance response of the fish. Previously we included a
likelihood component to incorporate this information in the
assessment model, but dropped it because this survey is now
modeled with a random walk in catchability, and a relatively
small systematic change in catchability is inconsequential
compared to other factors affecting catchability.

### Ageing error
An ageing error conversion matrix is used in the assessment model
to translate model population numbers at age to expected fishery
and survey catch at age (Table
1.\@ref(tab:age-error)). @dorn2003a estimated this matrix using
an ageing error model fit to the observed percent reader
agreement at ages 2 and 9. Mean percent agreement is close to
100% at age 1 and declines to 40% at age 10. Annual estimates of
percent agreement are variable, but show no obvious trend; hence
a single conversion matrix for all years in the assessment model
was adopted. The model is based on a linear increase in the
standard deviation of ageing error and the assumption that ageing
error is normally distributed. The model predicts percent
agreement by taking into account the probability that both
readers are correct, both readers are off by one year in the same
direction, and both readers are off by two years in the same
direction [@methot2000a]. The probability that both agree and
were off by more than two years was considered negligible. A
study evaluated pollock ageing criteria using radiometric methods
and found them to be unbiased [@kastelle2006a].

### Length frequency data
The assessment model was fit to length frequency data from
various sources by converting predicted age distributions (as
modified by age-specific selectivity) to predicted length
distributions using an age-length conversion matrix. This
approach was used only when age composition estimates were
unavailable, as occurs when the survey is the same as the
assessment. Because seasonal differences in pollock length at age
are large, particularly for the younger fish, several conversion
matrices were used. For each matrix, unbiased length
distributions at age were estimated for several years using
age-length keys, and then averaged across years. A conversion
matrix was estimated using 1992-1998 Shelikof Strait acoustic
survey data and used for winter survey length frequency data. The
following length bins were used: 5-16, 17-27, 28-35, 36-42,
43-50, 51-55, 56-70 (cm). Age data for the most recent
survey is now routinely available so this option does not need to
be invoked. A conversion matrix was estimated using second and
third trimester fishery age and length data during the years
(1989-1998), and was used when age composition data are
unavailable for the summer bottom trawl survey, which is only for
the most recent survey in the year that the survey is
conducted. The following length bins were used: 5-24, 25-34,
35-41, 42-45, 46-50, 51-55, 56-70 (cm), so that the
first four bins would capture most of the summer length
distribution of the age-1, age-2, age-3 and age-4 fish,
respectively. Bin definitions were different for the summer and
the winter conversion matrices to account for the seasonal growth
of the younger fish (ages 1-4).

### Initial data weighting
The input sample sizes were initially standardized by data set
before model tuning. Fishery age composition was given an initial
sample size of 200 except when the age sample in a given year
came from fewer than 200 hauls/deliveries, in which case the
number of hauls/deliveries was used. Both the Shelikof acoustic
survey and the bottom trawl age compositions were given an
initial sample size of 60, and the ADF&G crab/groundfish survey
was given a weight of 30.


## Parameters Estimated Outside the Assessment Model
Pollock life history characteristics, including natural
mortality, weight at age, and maturity at age, were estimated
independently outside the assessment model. These parameters are
used in the model to estimate spawning and population biomass and
obtain predictions of fishery catch and survey biomass. Pollock
life history parameters include:

- Natural mortality ($M$)
- Proportion mature at age
- Weight at age and year by fishery and by survey

### Natural mortality
@hollowed1990a estimated natural mortality ($M$) using a variety
of methods including estimates based on: a) growth parameters
[@alverson1975a; @pauly1980a], b) GSI [@gunderson1988a], c)
monitoring cohort abundance, and d) estimation in the assessment
model. These methods produced estimates of natural mortality that
ranged from 0.22 to 0.45. The maximum age observed was 22
years. Up until the 2014 assessment, natural mortality had been
assumed to be 0.3 for all ages.

@hollowed2000a developed a model for Gulf of Alaska pollock that
accounted for predation mortality. The model suggested that
natural mortality declines from 0.8 at age 2 to 0.4 at age 5, and
then remains relatively stable with increasing age. In addition,
stock size was higher when predation mortality was included. In a
simulation study, @clark1999a evaluated the effect of an
erroneous $M$ on both estimated abundance and target harvest
rates for a simple age-structured model. He found that "errors in
estimated abundance and target harvest rate were always in the
same direction, with the result that, in the short term,
extremely high exploitation rates can be recommended
(unintentionally) in cases where the natural mortality rate is
overestimated and historical exploitation rates in the
catch-at-age data are low." @clark1999a proposed that the chance
of this occurring could be reduced by using an estimate of
natural mortality on the lower end of the credible range, which
is the approach used in this assessment. In the 2014 assessment,
several methods to estimate of the age-specific pattern of
natural mortality were evaluated. Two general types of methods
were used, both of which are external to the assessment
model. The first type of method is based initially on theoretical
life history or ecological relationships that are then evaluated
using meta-analysis, resulting in an empirical equation that
relates natural mortality to some more easily measured quantity
such as length or weight. The second type of method is an
age-structured statistical analysis using a multispecies model or
single species model where predation is modeled. There are three
examples of such models for pollock in Gulf of Alaska, a single
species model with predation by @hollowed2000a, and two
multispecies models that included pollock by Kirk
[-@kirk2010a;-@kirk2012a]. These models were published in the
peer-reviewed literature, but likely did not receive the same
level of scrutiny as stock assessment models. Although these
models also estimate time-varying mortality, we averaged the
total mortality (residual natural mortality plus predation
mortality) for the last decade in the model to obtain a mean
age-specific pattern (in some cases omitting the final year when
estimates were much different than previous years). Use of the
last decade was an attempt to use estimates with the strongest
support from the data. Approaches for inclusion of time-varying
natural mortality will be considered in future pollock
assessments. The three theoretical/empirical methods used were
the following:


*@brodziak2011a*: Age-specific $M$ is given by

\begin{equation*}
 M(a)=
    \begin{cases}
      M_c \frac{L_{mat}}{L(a)} & \text{for } a<a_{mat}\\
      M_c & \text{for } a \geq a_{mat}
    \end{cases}
\end{equation*}

where $L_{mat}$ is the length at maturity, $M_c=0.30$ is the
natural mortality at $L_{mat}$, $L(a)$ is the mean length at age
for the summer bottom trawl survey for 1984-2013.

*@lorenzen1996a*: Age-specific $M$ for ocean ecosystems is given
by

$$ M(a)=\bar{W_a}^{-0.305}$$

where $\bar{W_a}$ is the mean weight at age from the summer
bottom trawl survey for 1984-2013.

*@gislason2010a*: Age-specific $M$ is given by

$$\ln(M)=0.55-1.61\ln(L)+1.44\ln(L_{\infty})+\ln{K}$$ where
$L_\infty = 65.2$ cm and $K = 0.30$ were estimated by fitting von
Bertalanffy growth curves using the NLS routine in R using summer
bottom trawl age data for 2005-2009 for sexes combined in the
central and western Gulf of Alaska.  Results were reasonably
consistent and suggest use of a higher mortality rate for age
classes younger than the age at maturity (Table 1.\@ref(tab:natM)
and Fig. 1.\@ref(fig:fig-natM)). Somewhat surprisingly, the
theoretical/empirical estimates were similar, on average, to
predation model estimates. To obtain an age-specific natural
mortality schedule for use in the stock assessment, we used an
ensemble approach and averaged the results for all methods. Then
we used the method recommended by Clay Porch in @brodziak2011a to
rescale the age-specific values so that the average for range of
ages equals a specified value. Age-specific values were rescaled
so that a natural mortality for fish greater than or equal to age
5, the age at 50% maturity, was equal to 0.3, the value of
natural mortality used in previous pollock assessments.

### Maturity at age
Maturity stages for female pollock describe a continuous process
of ovarian development between immature and post-spawning. For
the purposes of estimating a maturity vector (the proportion of
an age group that has been or will be reproductively active
during the year) for stock assessment, all fish greater than or
equal to a particular maturity stage are assumed to be mature,
while those less than that stage are assumed to be
immature. Maturity stages in which ovarian development had
progressed to the point where ova were distinctly visible were
assumed to be mature (i.e., stage 3 in the 5-stage pollock
maturity scale). Maturity stages are qualitative rather than
quantitative, so there is subjectivity in assigning stages, and a
potential for different technicians to apply criteria differently
(Neidetcher et al. 2014). Because the link between pre-spawning
maturity stages and eventual reproductive activity later in the
season is not well established, the division between mature and
immature stages is problematic. Changes in the timing of spawning
could also affect maturity at age estimates. @merati1993a
compared visual maturity stages with ovary histology and a blood
assay for vitellogenin and found general consistency between the
different approaches. @merati1993a noted that ovaries classified
as late developing stage (i.e., immature) may contain yolked
eggs, but it was unclear whether these fish would have spawned
later in the year. The average sample size of female pollock
maturity stage data per year since 2000 from winter acoustic
surveys in the Gulf of Alaska is 372 (Table 1.\@ref(tab:mat)).
In 2019, a new approach was introduced to estimate maturity at
age using specimen data from the Shelikof Strait acoustic
survey. Maturity estimates from 2003 onwards were revised using
this method. The approach uses local abundance to weight the
maturity data collected in a haul. To estimate abundance, each
acoustic survey distance unit (0.5 nmi of trackline) was assigned
to a stratum representing nearest survey haul. Each haul’s
biological data was then used to scale the corresponding acoustic
backscatter within that stratum into abundance. To generate
abundance weights for specimen data taken for each haul location,
the abundance estimates of adult pollock ($\geq 30$ cm fork
length) were summed for each haul-stratum. The 30 cm length
threshold represents the length at which pollock are 5% mature in
the entire Shelikof Strait historic survey data. Total adult
pollock abundances in each stratum was scaled by dividing by the mean
abundance per stratum (total abundance /number of
haul-strata). Weights range from 0.05 to 6, as some hauls were
placed in low-density regions while others sampled very dense
aggregations. For each haul, the number of female pollock
considered mature (prespawning, spawning, or spent) and immature
(immature or developing) were computed for each age. The maturity
ogive for maturity-at-age was estimated as a logistic regression
using a weighted generalized linear model where the dependent
variable was the binomial spawning state, the independent
variable was the age, and data from each haul were weighted by the
appropriate values as computed above. The length and age at 50%
maturity was derived (L50%, A50%) from the ratio of the
regression coefficients. The new maturity estimates had a
relatively minor impact on assessment results, and usually
reduced estimates of spawning biomass by about 2 percent.
Estimates of maturity at age in 2022 from winter acoustic surveys
using the new method are higher for younger fish, but lower for
older fish, compared to 2021 and the long-term mean for all ages
(Fig. 1.\@ref(fig:mat-all) and
1.\@ref(fig:mat-recent)). Inter-annual changes in maturity at age
may reflect environmental conditions, pollock population biology,
effect of strong year classes moving through the population, or
simply ageing error. Because there did not appear to be an
objective basis for excluding data, the 1983-2021 average
maturity at age was used in the assessment.

Logistic regression [@mccullagh1983a] was also used to
estimate the age and length at 50% maturity at age for each year
to evaluate long-term changes in maturation. Annual estimates of
age at 50% maturity are highly variable and range from 2.6 years
in 2017 to 6.1 years in 1991, with an average of 4.8 years
(Fig. 1.\@ref(fig:a50)). The last few years has shown a decrease in the age
at 50% mature, which is largely being driven by the maturation of
the 2012 year class at younger ages than is typical, however the
2019 to 2022 estimates of age at 50% mature are near the
long-term average. Length at 50% mature is less variable than the
age at 50% mature, suggesting that at least some of the
variability in the age at maturity can be attributed to changes
in length at age. Changes in year-class dominance also likely
affect estimates of maturity at length, as a similar pattern is
seen as with maturity at age with the 2012 cohort. The average
length at 50% mature for all years is approximately 43 cm.

### Weight at age
Year-specific fishery weight-at-age estimates are used in the
model to obtain expected catches in biomass. Where possible, year
and survey-specific weight-at-age estimates are used to obtain
expected survey biomass. For each data source, unbiased estimates
of length at age were obtained using year-specific age-length
keys. Bias-corrected parameters for the length-weight
relationship, $W=aL^b$ , were also estimated. Weights at age were
estimated by multiplying length at age by the predicted weight
based on the length-weight regressions. Weight at age for the
fishery, the Shelikof Strait acoustic survey, and the NMFS bottom
trawl survey and the summer acoustic survey are given in Tables
1.\@ref(tab:waa-fsh), 1.\@ref(tab:waa-srv1), and
1.\@ref(tab:waa-srv26). Data from the Shelikof Strait acoustic
survey indicates that there has been a substantial change in
weight at age for older pollock
(Fig. 1.\@ref(fig:fig-waa-srv1)). For pollock greater than age 6,
weight-at-age nearly doubled by 2012 compared to
1983-1990. However, weight at age trended strongly downward from
2012 to 2020, with some rebound in the last couple of
years. Further analyses are needed to evaluate whether these
changes are a density-dependent response to declining pollock
abundance, or whether they are environmentally forced. Changes in
weight-at-age have important implications for status
determination and harvest control rules.

A random effects (RE) model for weight at age [@ianelli2016a] was
used to estimate of fishery weight at age in `r year` since age
data were not available. The structural part of the model is an
underlying von Bertalanffy growth curve. Year and cohort effects
are estimated as random effects using the ADMB RE module. Further
details are provided in @ianelli2016a. Input data included
fishery weight age for 1975-`r year-1`. The model also
incorporates survey data by modeling an offset between fishery
and survey weight at age. Weight at age for the Shelikof Strait
acoustic survey (1981-`r tail(datfile$srv_acyrs1,1)`) and the
NMFS bottom trawl survey (1984-`r tail(datfile$srv_acyrs2,1)`)
were used. The model also requires input standard deviations for
the weight at age data, which are not available for GOA
pollock. In the 2016 assessment, a generalized variance function
was developed using a quadratic curve to match the mean standard
deviations at ages 3-10 for the eastern Bering Sea pollock
data. The standard deviation at age one was assumed to be equal
to the standard deviation at age 10. Survey weights at age were
assumed to have standard deviations that were 1.5 times the
fishery weights at age. A comparison of RE model estimates from
last year of the `r year-1` fishery weight at age with the data
now available indicate that the model overestimated weights
slightly (Fig. 1.\@ref(fig:fig-waa-fsh)). In this assessment, RE
model estimates of weight at age are used for the fishery in
`r year` and for yield projections and harvest recommendations.

@correa2023a details an exploratory and promising approach using
a state-space model to estimate the WAA within the assessment
model and this will be explored further in future years.

## Parameters Estimated Inside the Assessment Model
A large number of parameters are estimated when using this
modeling approach, though many are year-specific deviations in
fishery selectivity coefficients. Parameters were estimated using
Template Model Builder [TMB; @kristensen2016a], a modeling platform
based strongly on AD Model Builder [@fournier2012a] but which
contains improved functionality for estimating non-linear
hierarchical models. The model is determined to have converged
when the maximum parameter gradient is less than a small constant
(set to 1 x 10-6) and the Hessian matrix is invertible. Like AD Model
Builder, TMB includes post-convergence routines to calculate standard
errors (or likelihood profiles) for any quantity of interest, and
has state of the art Bayesian integration capabilities [@monnahan2018a].

A list of model parameters for the base model is shown below:

```{r}
read_xlsx(table.path, sheet='pars') %>%
 flextable()
```





