
# Harvest Recommendations

## Reference fishing mortality rates and spawning biomass levels

Since 1997, GOA pollock have been managed under Tier 3 of the
NPFMC tier system.  In Tier 3, reference mortality rates are
based on spawning biomass per recruit (SPR), while biomass
reference levels are estimated by multiplying the SPR by average
recruitment.  Estimates of the FSPR harvest rates were obtained
using the life history characteristics of GOA pollock (Table
1.\@ref(tab:spr-inputs)).  Spawning biomass reference levels were
based on mean 1978-`r year-1` age-1 recruitment (`r avgrec`
billion), which is `r pct.avgrec.change`% higher than the mean
value in last year's assessment. Spawning was assumed to occur on
March 15th, and a long-term average of maturity at age (1983-
`r year`) was used with mean spawning weight at age from the
Shelikof Strait acoustic surveys in 2017-`r year` to estimate
current reproductive potential.  Fishery weight at age was
assumed to be the most recent estimate from the RE model.
Pollock weight-at-age is highly variable, showing a sustained
increase, followed by a steep decline until a sharp increase in
2021 (Fig. 1.\@ref(fig:fig-waa-srv1)). The factors causing this
pattern are unclear, but are likely to involve both
density-dependent factors and environmental forcing. The SPR at
F=0 was estimated as `r round(spr0,3)` kg/recruit at age one.  FSPR rates
depend on the selectivity pattern of the fishery.  Selectivity
has changed as the fishery evolved from a foreign fishery
occurring along the shelf break to a domestic fishery on spawning
aggregations and in nearshore waters.  For SPR calculations,
selectivity was based on the average for `r year -4`-`r year` to
reflect current selectivity patterns. GOA pollock FSPR harvest
rates are given below:

```{r}
tab <- read_xlsx(table.path, sheet='SPR')
tab[,7] <- paste0(  format(round(100*dplyr::pull(tab,7),1), nsmall=1),"%")
flextable(tab) %>%
  colformat_double(digits=3, j=2) %>%
  colformat_double(digits=0, big.mark=',') 
```
<!-- The B40% estimate of `r b40` t represents a  -->
<!-- `r pct.b40.changeF`from the  B40% estimate of `r b40.last` t in the -->
<!-- `r year-1` assessment (Table 1.\@ref(tab:model-history)). The base model -->
<!-- projection of female spawning biomass in `r year+1` is  -->
<!-- `r ssb` t, which is `r pct.status` of unfished spawning biomass (based on -->
<!-- average post-1977 recruitment) and above B40% (`r b40` t), -->
<!-- thereby placing GOA pollock in sub-tier "a" of Tier 3.  -->

## 2022 acceptable biological catch

The definitions of OFL and maximum permissible FABC under
Amendment 56 provide a buffer between the overfishing level and
the intended harvest rate, as required by NMFS national standard
guidelines.  Since estimates of stock biomass from assessment
models are uncertain, the buffer between OFL and ABC provides a
margin of safety so that assessment error will not result in the
OFL being inadvertently exceeded. For GOA pollock, the maximum
permissible FABC harvest rate (i.e., FABC/FOFL) is 
`r format(100*fabc/fofl, digits=3)`% of the OFL harvest rate.
Projections for `r year+1` for the FOFL and the maximum
permissible FABC are given in Table 1.\@ref(tab:proj-scens).

### Should the ABC be reduced below the maximum permissible ABC?
The SSC in its December 2018 minutes recommended that all
assessment authors use the risk table when determining whether to
recommend an ABC lower than the maximum permissible.  The
following template is used to complete the risk table:

```{r}
 data.frame(level = c("Level 1: Normal",
                      "Level 2: Substantially increased concerns",
                      "Level 3: Major Concern",
                      "Level 4: Extreme concern"),
            assess = c("Typical to moderately increased uncertainty/minor unresolved issues in assessment.", 
                       "Substantially increased assessment uncertainty/ unresolved issues.",
                       "Major problems with the stock assessment; very poor fits to data; high level of uncertainty; strong retrospective bias.",
                       "Severe problems with the stock assessment; severe retrospective bias. Assessment considered unreliable."),
            pop = c("Stock trends are typical for the stock; recent recruitment is within normal range.",
                    "Stock trends are unusual; abundance increasing or decreasing faster than has been seen recently, or recruitment pattern is atypical.",
                    "Stock trends are highly unusual; very rapid changes in stock abundance, or highly atypical recruitment patterns.",
                    "Stock trends are unprecedented; More rapid changes in stock abundance than have ever been seen previously, or a very long stretch of poor recruitment compared to previous patterns."),
            enviro = c("No apparent environmental/ecosystem concerns",
                       "Some indicators showing adverse signals relevant to the stock but the pattern is not consistent across all indicators.",
                       "Multiple indicators showing consistent adverse signals a) across the same trophic level as the stock, and/or b) up or down trophic levels (i.e., predators and prey of the stock)",
                       "Extreme anomalies in multiple ecosystem indicators that are highly likely to impact the stock; Potential for cascading effects on other ecosystem components"),
            fish = c("No apparent fishery/resource-use performance and/or behavior concerns",
                     "Some indicators showing adverse signals but the pattern is not consistent across all indicators",
                     "Multiple indicators showing consistent adverse signals a) across different sectors, and/or b) different gear types",
                     "Extreme anomalies in multiple performance  indicators that are highly likely to impact the stock")) %>% 
   flextable() %>% 
     font(fontname = "Times", part = "all") %>% 
     width(j = ~ assess, width = 1.5) %>% 
     width(j = ~ pop, width = 2) %>% 
     width(j = ~ enviro, width = 1.5) %>% 
     width(j = ~ fish, width = 1.5) %>% 
     set_header_labels(level = "",
                     assess = "Assessment-related considerations",
                     pop = "Population dynamics considerations",
                     enviro = "Environmental/ecosystem considerations",
                     fish = "Fishery Performance") %>% 
   font(fontname = "Times", part = "all") %>% 
   fontsize(size = 10) %>% 
   theme_vanilla() %>% 
   italic(part = "header") 
```

The table is applied by evaluating the severity of four types of
considerations that could be used to support a scientific
recommendation to reduce the ABC from the maximum permissible.
These considerations are stock assessment considerations,
population dynamics considerations, environmental/ecosystem
considerations, and fishery performance.  Examples of the types
of concerns that might be relevant include the following:

1.  Assessment considerations---data-inputs: biased ages,
    skipped surveys, lack of fishery-independent trend data;
    model fits: poor fits to fits to fishery or survey data,
    inability to simultaneously fit multiple data inputs; model
    performance: poor model convergence, multiple minima in the
    likelihood surface, parameters hitting bounds; estimation
    uncertainty: poorly-estimated but influential year classes;
    retrospective bias in biomass estimates.
2.  Population dynamics considerations---decreasing biomass
    trend, poor recent recruitment, inability of the stock to
    rebuild, abrupt increase or decrease in stock abundance.
3.  Environmental/ecosystem considerations---adverse trends in
    environmental/ecosystem indicators, ecosystem model results,
    decreases in ecosystem productivity, decreases in prey
    abundance or availability, increases or increases in predator
    abundance or productivity.
4.  Fishery performance---fishery CPUE is showing a contrasting
    pattern from the stock biomass trend, unusual spatial pattern
    of fishing, changes in the percent of TAC taken, changes in
    the duration of fishery openings.

#### Assessment considerations

A continuing assessment issue is the conflicting size of the 2018
year class abundance between the sources of information.  The
winter pre-spawning Shelikof acoustic survey 2019 estimate was
indicative of a strong year class, but the 2020 estimate of age 2
fish was only 10% of the long-term average.  Over the full
Shelikof time series, high age-1 estimates have always
been followed by high age-2 estimates in the next year
(Fig. 1.9).  It was previously hypothesized that the 2018 year
class could have moved out of Shelikof Strait or experienced
unusually high mortality. In 2021, both the 2020 age 2 ADF&G and
2021 age 3 Shelikof survey observed proportions were low relative
to the model expectation, providing further evidence of a reduced
2018 cohort. In contrast, the 2020 fishery catch at age was very
close to expected for age-2 fish, and there are some apparent
age-3 fish in the 2021 length compositions from the summer
surveys.

This year brings important new data sets to the table, which help
to resolve this uncertainty. Shelikof age compositions provide
information on the 2018 cohort as age-4s, while the 2021 fishery,
NMFS bottom trawl and summer acoustic surveys provide information
age-3s. By plotting Pearson residuals by cohort, it is
immediately clear that the Shelikof survey is the main source of
information suggesting a small cohort (i.e., negative residuals),
while the other surveys and fishery generally suggest a larger
cohort.

```{r}
addfig('resid_by_cohort.png')
```

The estimated size of the cohort changed little between the 2021
and 2022 models, suggesting that the estimate likely has settled
down at 5.45 billion (CV=18%). This cohort does not appear
unusual compared to other recent cohort estimates as data are
added.

```{r}
addfig('retro_recruits.png')
```

In summary, the 2018 cohort is slightly larger than average, and
the initial Shelikof estimate for age-1s was anomalously high,
while all subseqeunt years were low. The other surveys and
fishery appear to be in line with this estimate, building a
consensus of the result. The one notable exception is the the
2020 ADF&G estimate which also had a large negative age-2
residual. Since this survey targets larger, older fish it is not
compelling evidence, but it will be interesting to see how the
age-4 residual turns out in the 2023 assessment when those data
are available. The implication is that the 2018 cohort
availability to the Shelikof survey has been low for the last
three years after an anomolously high age-1 estimate. It is not
apparent why this would be, and appears to have no historical
precedent. The cohort maybe have been distributed in a different
area as immature age 2-3 fish, with the cohort's availability to
the Shelikof survey increasing as a larger proportion of the
cohort reaches maturity, but since surveys for other winter
spawning areas have been limited in recent years this can only
be speculated. Future surveys of other spawning areas could prove
valuable to help resolve this mystery. It could also be a
statistical cooincidence.

Despite the unusual characteristics of this cohort there are no
assessment model concerns, as cohort size is estimated
consistently and other data sources are in agreement and the
model fits them well. We thus gave assessment considerations a
score of 1 --- no increased concerns.

#### Population dynamics considerations

The large 2012 year class had a strong impact on the recent
pollock population, from a steep decline in age diversity
(Fig. 1.\@ref(fig:fsh-chars)) to abnormal growth and maturation
(but not mortality as previously suspected;
Fig. 1.\@ref(fig:2012-bio)), which had led to an increase in
concern. However, this year class is no longer the predominant
one in the fishery and two large ones (2017 and 2018) have
already entered the fishery, with another large one in 2020 to
enter in the coming years (Figs. 1.\@ref(fig:ages-fsh) and
1.\@ref(fig:ssb-status-rec)), resulting in a return to normal age
diversity since 2020. Consequently, we gave populations
dynamics considerations a score of 1---no increased concerns.

#### Environmental/Ecosystem considerations
Appendix 1A provides a detailed look at environmental/ecosystem
considerations specific to this stock within the ecosystem and
socioeconomic profile (ESP). Broad-scale information on
environmental and ecosystem considerations are provided by the
Gulf of Alaska Ecosystem Status Report (GOA ESR;
@ferriss2022a). The text below summarizes ecosystem information
related to GOA pollock provided from both the ESP and GOA ESR.

**Environmental Processes:** The 2022 and predicted 2023 ocean
temperatures are all within known optimal ranges for pollock life
history stages (spawning 150-300m: 1-7°C, egg 0-200m, 5-6°C,
larva surface 3-7°C, as referenced in the ESP). It is reasonable
to expect that the 2022 and predicted 2023 average ocean
temperatures will provide good spawning habitat, but warmer fall
surface temperatures may negatively impact growth to a size that
promotes over-winter survival for this year’s age class. Spring
surface temperatures over the western and central GOA were
slightly above average (Appendix 1A: Spring Temperature Surface
WCGOA Satellite indicator by M. Callahan). Western GOA surface
temperatures were cooler than average in the winter (Satellite,
@lemagie2022a), transitioned from cool to above average in spring
(Satellite, @lemagie2022a, 5.7°C Seward Line, @danielson2022a),
above average in summer (12.3°C Seward Line, @danielson2022a and
Satellite, @lemagie2022a) and fall (Satellite,
@lemagie2022a). To-date, the western GOA is experiencing the 3rd
consecutive non-marine heatwave year, with one short-term
heatwave event in July (Satellite, @lemagie2022a); however, the
central GOA experienced average heatwave events this year which
is an increase from last year (Appendix 1A: Annual Heatwave GOA
Model indicator by S. Barbeaux). Spring winds from the northeast
in Shelikof Strait (downwelling favorable, flowing south through
Shelikof Strait) were downwelling favorable, contributing to
retention of 2022 larvae and potential for a stronger age-1 year
class in 2023, similar to conditions in 2021, 2020, and 2012
(Appendix 1A: Spring Wind Direction Kodiak Buoy by
L. Rogers). Over the western and central GOA, spring chlolorphyll
*a* concentration increased to slightly below average while the
peak was slightly earlier than last year by still later than
average (Appendix 1A: Spring Chlorophylla Biomass and Peak WCGOA
Satellite by M. Callahan). For the western GOA, spring primary
productivity varied spatially from below to above average
chlorophyll a concentrations, with slightly later than average
spring bloom timing (Satellite, @gann2022a). Elevated spring
productivity was observed along the Seward Line (CGOA), in terms
of a high phytoplankton size index, inferring increased energy
transfer from the base of the food web [@strom2022a]. Upcoming
2023 winter and spring surface temperatures are predicted to be
cooler than average, in alignment with La Nina conditions and a
negative Pacific Decadal Oscillation.

**Prey**:
Planktivorous foraging conditions were potentially above average
across the GOA in 2022 with limited data in western GOA. Moderate
and regionally variable across the GOA in 2021. The biomass of
large calanoid copepods was average to above average at the
eastern edge of western GOA (Seward Line, @hopcroft2022a). Small
and large calanoid copepods increased to above average, while
euphausiid biomass decreased to approximately average in eastern
GOA inside waters (AFSC SECM Survey, Icy Strait, Fergusson
2022). Planktivorous seabird reproductive success, an indicator
of zooplankton availability and nutritional quality, was below
average just south of Kodiak (Chowiet Island), and above average
in central GOA (Middleton Island on shelf edge off Seward) and
eastern GOA (St. Lazaria Island near Sitka; @drummond2022a; Hatch
2022, Appendix 1A: Annual Auklet Reproductive Success Chowiet
Survey indicator by S. Zador). Catch-per-unit-effort of age-0
pollock decreased to just above average in the nearshore beach
seine survey around Kodiak and along the Alaska Peninsula
suggesting less productive feeding conditions in the nearshore.

**Predators and Competitors**: Predation pressure from key
groundfish species (arrowtooth flounder, Pacific cod, Pacific
halibut, and potentially sablefish) is expected to be
moderate. Pacific cod and halibut biomass have remained
relatively low, arrowtooth flounder assessment shows slight
decline in 2022, (and a slight increase in 2022 ADF&G
survey). The sablefish assessment shows a slight decline and the
large 2016 age class of sablefish will have moved to adult slope
habitat with little overlap with pollock, however a larger 2019
year class of sablefish may bring increased predation pressure
(Appendix 1A: Annual Sablefish Biomass GOA Model by
K. Shotwell). Western GOA Steller sea lions slightly increased
but remain lower than previous biomass peaks (Appendix 1A: Annual
Steller Sea Lion Adult GOA Survey by K. Sweeney). Cannibalism is
less prevalent in the GOA as in other Alaskan waters. Potential
competitors are a large year class of juvenile sablefish (2019)
as well as other potential strong sablefish year classes, and a
relatively large population of Pacific Ocean perch (Appendix 1A:
Annual Sablefish and Pacific Ocean Perch Biomass GOA Model by
K. Shotwell). Pink salmon contributed lower competitive pressure
in 2022 given their predictably lower returns in even years
[@whitehouse2022a;@vulstek2022a].

*Summary for Environmental/Ecosystem Considerations*

* Environment: Moderate thermal conditions for adults and larvae,
with concerns of potential impacts of fall surface warmth on YOY
winter survival.

*Prey: Planktivorous and piscivorous foraging conditions were
above average in the WGOA and EGOA, with high uncertainty due to
limited zooplankton data (especially euphausiid data which are
primary prey for juveniles & adults). The age-0 year classes
sampled in the beach seine survey decreased to slightly above
average suggesting slightly limited
prey resources.

*Competitors/Predators: Predation pressure is expected to be
moderate from key groundfish species (arrowtooth flounder,
Pacific cod, Pacific halibut, and potentially sablefish) and
potential competitors for zooplankton may remain moderate given a
large year class of juvenile sablefish (2019), and large
population of Pacific Ocean perch but lower (even year) returns
of pink salmon.

Currently the 2018, 2017 and 2012 year classes are the dominant
year classes supporting the fishery, and there is no cause to
suspect unfavorable conditions for those cohorts. Together, the
most recent data available suggest an ecosystem risk Level 1---
Normal: No apparent environmental/ecosystem concerns.


#### Fishery performance
Trends in effort-weighted fishery CPUE were examined in the ESP
(Appendix 1A) for two seasons, the pre-spawning fishery (A and B
seasons) and the summer/fall fishery (C and D seasons). Fishery
CPUE is either above (A and B seasons) or close to (C and D
seasons) the long-term average, and is very consistent with the
abundance trend of exploitable biomass from the assessment. No
concerns regarding fishery performance were identified and this
element was given a score of 1.


#### Summary and ABC recommendation

```{r}
l1 = "Level 1: No increased concerns"
l2 = "Level 2: Substantially increased concerns"
l3 = "Level 3: Major concern"
l4 = "Level 4: Extreme concern"

data.frame(assess = l1,
           pop = l1,
           enviro = l1,
           fish = l1) %>%
  flextable() %>%
    width(j = ~ assess, width = 1.5) %>%
    width(j = ~ pop, width = 1.5) %>%
    width(j = ~ enviro, width = 1.5) %>%
    width(j = ~ fish, width = 1.5) %>%
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>%
  fontsize(size = 10) %>%
  font(fontname = "Times", part = "all") %>%
  italic(part = "header")
```

Given the lack of elevated scores in the risk table, the author’s
recommended ABC is based on the maximum permissible ABC,
resulting in a `r year+1` ABC of `r maxabc` t, which is a 
`r pct.abc.changeF`from the `r year` ABC. The author’s recommended
`r year+2` ABC is `r F(exec_table[12,4])` t. The OFL in 
`r year+1` is `r ofl` t, and the OFL in `r year+2` if the ABC is
taken in `r year+1` is `r F(exec_table[10,4])` t.  

To evaluate the probability that the stock will drop below the
B20% threshold, we projected the stock forward for five years
using the author’s recommended fishing mortality schedule. This
projection incorporates uncertainty in stock status, uncertainty
in the estimate of B20%, and variability in future
recruitment. We then sampled from the probability of future
spawning biomass using Markov chain Monte Carlo (MCMC) using the
no-U-turn sampler available in ADMB [@monnahan2018a]. Analysis of
the posterior samples indicates that probability of the stock
dropping below B20% will be negligible through `r year+5`,
conditional upon the model specified here
(Fig. 1.\@ref(fig:depletion)).

### Projections and Status Determination
<!-- !!!! this stuff came from the template, but Martin's SAFE -->
<!-- didn't have it.. should probably add later

<!-- Under the MSFCMA, the Secretary of Commerce is required to report -->
<!-- on the status of each U.S. fishery with respect to overfishing. -->
<!-- This report involves the answers to three questions: -->

<!-- 1. Is the stock being subjected to overfishing? -->
<!-- 2. Is the stock currently overfished? -->
<!-- 3. Is the stock approaching an overfished condition? -->

<!-- *Is the stock being subjected to overfishing?* The official catch -->
<!-- estimate for the most recent complete year (`r year-1`) is -->
<!-- *correct this later* r catch %\>% filter(Year==year-1) %\>% -->
<!-- pull(Catch) %\>% format(., big.mark = ",")`t.  This is less than -->
<!-- the`r year-1` OFL of 5,402 t.  Therefore, the stock is not being -->
<!-- subjected to overfishing. -->

<!-- Harvest Scenarios #6 and #7 are intended to permit determination -->
<!-- of the status of a stock with respect to its minimum stock size -->
<!-- threshold (MSST).  Any stock that is below its MSST is defined to -->
<!-- be overfished.  Any stock that is expected to fall below its MSST -->
<!-- in the next two years is defined to be approaching an overfished -->
<!-- condition.  Harvest Scenarios #6 and #7 are used in these -->
<!-- determinations as follows: -->

<!-- *Is the stock currently overfished?* This depends on the stock's estimated spawning biomass in `r year`: -->
<!-- -  If spawning biomass for `r year` is estimated to be below ½ -->
<!--    $B_{35\%}$, the stock is below its MSST. -->
<!-- -  If spawning biomass for `r year` is estimated to be above -->
<!--    $B_{35\%}$ the stock is above its MSST. -->
<!-- -  If spawning biomass for `r year` is estimated to be above ½ -->
<!-- $B_{35\%}$ but below $B_{35\%}$, the stock's status relative to -->
<!-- MSST is determined by referring to harvest Scenario #6 (Table -->
<!-- 10.16). If the mean spawning biomass for 2028 is below -->
<!-- $B_{35\%}$, the stock is below its MSST. Otherwise, the stock is -->
<!-- above its MSST.   -->

<!-- *Is the stock approaching an overfished -->
<!-- condition?* This is determined by referring to harvest Scenario -->
<!-- #7: -->
<!-- -  If the mean spawning biomass for `r year+2` is below 1/2 -->
<!--    $B_{35\%}$, the stock is approaching an overfished condition. -->
<!-- -  If the mean spawning biomass for `r year+2` is above -->
<!--    $B_{35\%}$, the stock is not approaching an overfished -->
<!--    condition. -->
<!-- -  If the mean spawning biomass for `r year+2` is above 1/2 -->
<!--    $B_{35\%}$ but below $B_{35\%}$, the determination depends on -->
<!--    the mean spawning biomass for `r year+12` If the mean spawning -->
<!--    biomass for `r year+12` is below $B_{35\%}$, the stock is -->
<!--    approaching an overfished condition. Otherwise, the stock is -->
<!--    not approaching an overfished condition. Based on the above -->
<!--    criteria and Table 10.16, the stock is not overfished and is -->
<!--    not approaching an overfished condition. -->

<!-- The fishing mortality that would have produced a catch for last year equal to last year's OFL is `r best_f`. -->

A standard set of projections is required for each stock managed
under Tiers 1, 2, or 3 of Amendment 56.  This set of projections
encompasses seven harvest scenarios designed to satisfy the
requirements of Amendment 56, the National Environmental Policy
Act, and the Magnuson-Stevens Fishery Conservation and Management
Act (MSFCMA). For each scenario, the projections begin with the
vector of `r year` numbers at age as estimated in the assessment.
This vector is then projected forward to the beginning of 
`r year + 1` using the schedules of natural mortality and
selectivity described in the assessment and the best available
estimate of total (year-end) catch for `r year`.  In each
subsequent year, the fishing mortality rate is prescribed on the
basis of the spawning biomass in that year and the respective
harvest scenario.  In each year, recruitment is drawn from an
inverse Gaussian distribution whose parameters consist of maximum
likelihood estimates determined from recruitments estimated in
the assessment.  Spawning biomass is computed in each year based
on the time of peak spawning and the maturity and weight
schedules described in the assessment.  Total catch after
`r year` is assumed to equal the catch associated with the
respective harvest scenario in all years. This projection scheme
is run 1,000 times to obtain distributions of possible future
stock sizes, fishing mortality rates, and catches.

Five of the seven standard scenarios will be used in an
Environmental Assessment prepared in conjunction with the final
SAFE.  These five scenarios, which are designed to provide a
range of harvest alternatives that are likely to bracket the
final TAC for `r year+1`, are as follow ($\text{max}F_{ABC}$ refers
to the maximum permissible value of $F_{ABC}$ under Amendment
56):

- Scenario 1: In all future years, *F* is set equal to $max
  F_{ABC}$. (Rationale: Historically, TAC has been constrained
  by ABC, so this scenario provides a likely upper limit on
  future TACs.)

- Scenario 2: In `r year` and `r year + 1`, *F* is set equal to a
  constant fraction of $max F_{ABC}$, where this fraction is
  equal to the ratio of the realized catches in `r year-3`-
  `r year-1` to the ABC recommended in the assessment for each of
  those years.  For the remainder of the future years, maximum
  permissible ABC is used.  (Rationale: In many fisheries the ABC
  is routinely not fully utilized, so assuming an average ratio
  catch to ABC will yield more realistic projections.)

- Scenario 3: In all future years, *F* is set equal to 50% of
    $max F_{ABC}$. (Rationale: This scenario provides a likely
    lower bound on FABC that still allows future harvest rates to
    be adjusted downward when stocks fall below reference
    levels.)

- Scenario 4: In all future years, *F* is set equal to the 
  `r year-4`-`r year` average *F*. (Rationale: For some stocks, TAC
  can be well below ABC, and recent average *F* may provide a
  better indicator of $F_{TAC}$ than $F_{ABC}$.)

- Scenario 5: In all future years, *F* is set equal to zero.
  (Rationale: In extreme cases, TAC may be set at a level close
  to zero.)

Two other scenarios are needed to satisfy the MSFCMA's
requirement to determine whether a stock is currently in an
overfished condition or is approaching an overfished condition.
These two scenarios are as follows (for Tier 3 stocks, the MSY
level is defined as $B_{35\%}$):

- Scenario 6: In all future years, *F* is set equal to $F_{OFL}$.
  Rationale: This scenario determines whether a stock is
  overfished. If the stock is expected to be 1) above its MSY
  level in `r year` or 2) above 1/2 of its MSY level in 
  `r year` and above its MSY level in `r year+9` under this scenario, then
  the stock is not overfished.

- Scenario 7: In `r year+1` and `r year + 2`, *F* is set equal to
  max $F_{ABC}$, and in all subsequent years F is set equal to
  FOFL.  Rationale: This scenario determines whether a stock is
  approaching an overfished condition. If the stock is 1) above
  its MSY level in `r year+2` or 2) above 1/2 of its MSY level in
  `r year+2` and expected to be above its MSY level in
  `r year+11` under this scenario, then the stock is not approaching
  an overfished condition.

Results from scenarios 1-7 are presented in Table
1.\@ref(tab:proj-scens). Mean spawning biomass is projected to
decline to 2026 under full exploitation scenarios, but will
increase under the F=0 and other low exploitation scenarios
(Fig. 1.\@ref(fig:fig-proj-scens)). We project catches to
increase through 2024, and then drop slightly in subsequent
years.

Under the MSFCMA, the Secretary of Commerce is required to report
on the status of each U.S. fishery with respect to
overfishing. This report involves the answers to three questions:
1) Is the stock being subjected to overfishing? 2) Is the stock
currently overfished? 3) Is the stock approaching an overfished
condition?

The catch estimate for the most recent complete year (`r year-1`)
is `r F(tail(datfile$cattot,2)[1])` t, which is less than the
`r year-1` OFL of `r ofl.last` t. Therefore, the stock is not
subject to overfishing. The fishing mortality that would have
produced a catch in `r year-1` equal to the `r year-1` OFL is
0.233.

Scenarios 6 and 7 are used to make the MSFCMA’s other required
status determination as follows:

Under scenario 6, spawning biomass is estimated to be `r ssb` t
in `r year` (see Table 1.\@ref(tab:proj-scens)), which is above
B35% (`r b35` t). Therefore, GOA pollock is not currently
overfished.

Under scenario 7, projected mean spawning biomass in `r year +2`
is `r F(exec_table[3,4])` t, which is above B35% (`r b35`
t). Therefore, GOA pollock is not approaching an overfished
condition.


The recommended area apportionment to management areas in the
central and western portions of the Gulf of Alaska
(central/western/west Yakutat) are provided in Appendix 1D.
