### Should the ABC be reduced below the maximum permissible ABC?
The SSC in its December 2018 minutes recommended that all
assessment authors use the risk table when determining whether to
recommend an ABC lower than the maximum permissible.  The
following template is used to complete the risk table, which was
updated in 2023 to reflect only three levels of concern:

```{r}
 data.frame(level = c("Level 1: No Concern",
                      "Level 2: Major Concern",
                      "Level 3: Extreme concern"),
            assess = c("Typical to moderately increased uncertainty/minor unresolved issues in assessment.", 
                       "Major problems with the stock assessment; very poor fits to data; high level of uncertainty; strong retrospective bias.",
                       "Severe problems with the stock assessment; severe retrospective bias. Assessment considered unreliable."),
            pop = c("Stock trends are typical for the stock; recent recruitment is within normal range.",
                    "Stock trends are highly unusual; very rapid changes in stock abundance, or highly atypical recruitment patterns.",
                    "Stock trends are unprecedented; More rapid changes in stock abundance than have ever been seen previously, or a very long stretch of poor recruitment compared to previous patterns."),
            enviro = c("No apparent environmental/ecosystem concerns",
                       "Multiple indicators showing consistent adverse signals a) across the same trophic level as the stock, and/or b) up or down trophic levels (i.e., predators and prey of the stock)",
                       "Extreme anomalies in multiple ecosystem indicators that are highly likely to impact the stock; Potential for cascading effects on other ecosystem components"),
            fish = c("No apparent fishery/resource-use performance and/or behavior concerns",
                     "Multiple indicators showing consistent adverse signals a) across different sectors, and/or b) different gear types",
                     "Extreme anomalies in multiple performance  indicators that are highly likely to impact the stock")) %>% 
   flextable() %>% 
     font(fontname = "Times", part = "all") %>% 
     width(j = ~ assess, width = 1.5) %>% 
     width(j = ~ pop, width = 2) %>% 
     width(j = ~ enviro, width = 1.5) %>% 
     width(j = ~ fish, width = 1.5) %>% 
     set_header_labels(level = "",
                     assess = "Assessment-related considerations",
                     pop = "Population dynamics considerations",
                     enviro = "Environmental/ecosystem considerations",
                     fish = "Fishery Performance") %>% 
   font(fontname = "Times", part = "all") %>% 
   fontsize(size = 10) %>% 
   theme_vanilla() %>% 
   italic(part = "header") 
```

The table is applied by evaluating the severity of four types of
considerations that could be used to support a scientific
recommendation to reduce the ABC from the maximum permissible.
These considerations are stock assessment considerations,
population dynamics considerations, environmental/ecosystem
considerations, and fishery performance.  Examples of the types
of concerns that might be relevant include the following:

1.  Assessment considerations---data-inputs: biased ages,
    skipped surveys, lack of fishery-independent trend data;
    model fits: poor fits to fits to fishery or survey data,
    inability to simultaneously fit multiple data inputs; model
    performance: poor model convergence, multiple minima in the
    likelihood surface, parameters hitting bounds; estimation
    uncertainty: poorly-estimated but influential year classes;
    retrospective bias in biomass estimates.
2.  Population dynamics considerations---decreasing biomass
    trend, poor recent recruitment, inability of the stock to
    rebuild, abrupt increase or decrease in stock abundance.
3.  Environmental/ecosystem considerations---adverse trends in
    environmental/ecosystem indicators, ecosystem model results,
    decreases in ecosystem productivity, decreases in prey
    abundance or availability, increases or increases in predator
    abundance or productivity.
4.  Fishery performance---fishery CPUE is showing a contrasting
    pattern from the stock biomass trend, unusual spatial pattern
    of fishing, changes in the percent of TAC taken, changes in
    the duration of fishery openings.

#### Assessment considerations

Several important assessment considerations arose in 2023. The
new abundance indices had conflict, with two going up and two
going down. The two surveys which cover the whole extent of the
Gulf were both up more than 70% from two years ago, while the two
with more limited spatial coverage were down from 2022. Not
surprisingly the model was not able to fit these data points, and
in particular the NMFS bottom trawl index has not fit well since
2015 (Fig. 1.\@ref(fig:index-fits)). 


The Shelikof abundance estimate is unexpectedly low and comprised
predominantly of pollock greater than 40 cm, but also coincides
with increased estimates of biomass in outlying areas, like the
Chirikof shelf break and Shumagin islands, relative to recent
years (Fig. 1.\@ref(fig:biomass-area-srv1)). However, biomass
estimates in the outlying areas are not abnormally high compared
to historical surveys and the Shelikof estimate still constitutes
the largest winter spawning area by far. This low Shelikof
estimate is thus not well explained by spatial shifts. A mismatch
between spawn timing and survey timing is another reasonable
hypothesis, as the two were linked statistically via a
relationship with catchability as was presented to the Groundfish
Plan Team in September 2022. However, the covariate values in
2023 would suggest only moderately lower catchability and likely
also not explain the poor fit to the new data point. In the end,
the fit is poor but not unprecedented with this stock, and the
previous few years have fit very well. We therefore believe there
is no reason for a substantial concern.

In contrast, the NMFS bottom trawl survey has fit poorly for the
last 5 biennial surveys (since 2015). The expected trend is the
opposite of the observed trend. The consistency of this misfit is
a larger concern, particularly because the prior on catchability
for this index is an important contributor to estimating the
scale of the stock. We therefore consider this an elevated stock
assessment concern.

Finally, we highlight the poor retrospective pattern estimated
this year. The estimated $\rho=-0.308$ is considered significant,
with substantial and consistent increases of estimated SSB as
data are added to the model. We hypothesized that the inflexible
time-varying catchability for the winter Shelikof survey, coupled
with the notably low estimate in 2023, could exaggerate this
retrospective pattern. To test this hypothesis we estimated the
process error for the random walk (assumed to be 0.05 in the base
model) to allow it increased flexibility. This type of estimation
is possible because model 23.0 is in TMB. The estimated value of
the process error was 0.36, a substantial increase as expected,
but the retrospective pattern was actually worse and decreased to
-0.40 (results not shown). It thus appears that this
retrospective pattern is not driven by an inadequately flexible
catchability for the Shelikof survey. Further explorations of
causes and reasonable solutions to minimize this significant
retrospective pattern will be done, but for now we highlight it
as a stock assessment concern.

Between the poor model fits and significant retrospective pattern
we assign level 2: major concerns with assessment
considerations. 

#### Population dynamics considerations

The large 2012 year class has had a strong impact on the recent
pollock population, from a steep decline in age diversity
(Fig. 1.\@ref(fig:fsh-chars)) to abnormal growth and maturation
(but not mortality as previously suspected), which had led to an
increase in concern. The estimated size of this cohort has
increased substantially over the last few years, including a 9.2%
and 9.9% increase in 2022 and 2023 with additional data, an
increase of almost 10 billion fish from 2020 to 2023. For
context, this increase alone would be considered a large cohort,
and only magnifies the large impact it has on the population
dynamics over the last 10 years. However, the 2012 year class,
now in the 10+ age class, is no longer the predominant one in the
fishery and two large ones (2017 and 2018) have already entered
the fishery, with another large one in 2020 to enter in the
coming years (Figs. 1.\@ref(fig:ages-fsh) and
1.\@ref(fig:ssb-status-rec)), resulting in a return to normal age
diversity and population dynamics.

A new phenomenon emerged in the last handful of years that is
worth highlighting. Many of the estimates of recent cohort sizes
are abnormally small compared to previous estimates as seen in
the following figure.

```{r} 
 addfig("log_recruits.png")
```

These recruits have implications for $\sigma_R$. We used TMB's
ability to estimate process errors during a retrospective peel to
test how sensitive the quantity is to these cohorts (below). We
found a big increase from 2015 to 2016, and again from 2022
to 2023. This demonstrates that these small cohorts have a strong
influence on the perception of the variation in recruitment. A
value of $\sigma_R$ closer to 1.8 is estimated by the model using
data through 2023, but $\sigma_R=1.3$ is assumed in the model
this year. Future work will be done to further corroborate the
small cohorts and explore whether recruitment variation should be
updated. For now we consider 1.3 a more reasonable value.

```{r}
 addfig("sigmaR_ests.png")
```

These vanishingly small cohorts are clear aberrations and violate
the statistical assumption of the assessment, namely that
log-recruitment is normally distributed. They also are
consistently estimated (Fig. 1\@ref(fig:retro-recruits)) and
corroborated by larval surveys. In other words, the estimates
appear justified by the data. Whether this shift is caused by
environmental forcing or other factors is unclear. However, from
a population dynamics perspective, which considers recruitment in
natural space, the effects are expected to be minimal. This is
because the cohorts are not expected to contribute to the
spawning stock whether they are very small or exceptionally
small. It is the large cohorts that drive the population dynamics
in the end. So while we highlight these abnormal recruitment
failures, we do not believe it rises to the level of a major
concern. We therefore give a level 1: no concern to population
dynamics considerations.

#### Environmental/Ecosystem considerations
Appendix 1A provides a detailed look at environmental/ecosystem
considerations specific to this stock within the ecosystem and
socioeconomic profile (ESP). Broad-scale information on
environmental and ecosystem considerations are provided by the
Gulf of Alaska Ecosystem Status Report (GOA ESR;
@ferriss2022a). The text below summarizes ecosystem information
related to GOA pollock provided from both the ESP and GOA ESR.

**Environmental Processes:** The 2023 ocean temperatures are all
within known optimal ranges for pollock life history stages
(Feb-May 150-300m: spawning 1-7°C, Mar-Apr 0-200m: egg 5-6°C,
Apr-Jul surface: larva 3-7°C, as referenced in the ESP, Appendix
1A). Western GOA temperatures at depth on the shelf were
approximately average. Surface waters were approximately average/
cooler than average in the winter, spring, and fall with warmer
waters in the summer (Satellite: @lemagie2023a; Appendix 1A:
Callahan, Seward Line: @danielson2023a, NOAA bottom trawl:
@o2023a, Shelikof: @axler2023a). The central GOA experienced
below average marine heatwave events this year, a decrease from
last year (Appendix 1A: Barbeaux). The mean direction of the
spring wind was southwest down Shelikof Strait suggesting
retention in suitable larval habitat but sustained April offshore
gap winds near Kodiak may have altered advective patterns
(Appendix 1A: Rogers). Over the western and central GOA, spring
chlorophyll-a concentrations were below average and the peak
spring bloom was considerably late (Satellite, Appendix 1A:
Callahan). Upcoming 2024 winter and spring surface temperatures
are predicted to be warmer than average, in alignment with El
Niño conditions, potentially impacting larval pollock survival
(depending on intensity and duration of the warming event).

**Prey**: Zooplankton biomass were below average to average on
the GOA shelf in the spring and summer. Zooplankton biomass in
the WGOA progressed from below average in the spring (lower small
and calanoid copepod biomass and higher euphausiid biomass) to
improved conditions in the summer (above average biomass of large
calanoid copepods and euphausiids, but continued lower small
copepod biomass; Appendix 1A: Rogers, Shelikof St., and Seward
Line, @hopcroft2023a). Summer planktivorous foraging conditions
were somewhat improved with above average large calanoid copepod
and euphausiid biomass, but continued lower small copepod biomass
(Shelikof, @kimmel2023a). Planktivorous seabird reproductive
success, an indicator of zooplankton availability and nutritional
quality, was approximately average just south of Kodiak (Chowiet
Island), and in the central GOA (Middleton Island)
(@drummond2023a, @whelan2023a, Appendix 1A). Adult and juvenile
fish conditions were below average (Bottom trawl survey,
@o2023a; winter acoustic survey, Appendix 1A:
Monnahan). Percent euphausiids in the diet of juveniles was
slightly above average (Appendix 1A: Aydin). Catches of larval
and YOY pollock in spring and summer surveys were low (Shelikof
St, @rogers2023a, Appendix 1A: Rogers, Kodiak beach seine survey,
Appendix 1A: Laurel), suggesting less productive feeding
conditions in the nearshore for larval pollock.

**Predators and Competitors**: Predation pressure from key
groundfish species (arrowtooth flounder, Pacific cod, Pacific
halibut, and potentially sablefish) is expected to be
moderate. Pacific cod, P. halibut, and arrowtooth flounder
biomass have remained relatively low (@hulson2023a,
@whorton2023a, Appendix 1A: Shotwell). The sablefish population
has had multiple large age classes since 2016, potentially adding
predation pressure to pollock prior to moving to adult slope
habitat (sablefish assessment, @goethel2023a, Appendix
1A). Western GOA Steller sea lions were not reassessed in 2023
but remain lower than previous biomass peaks
(@sweeney2022a). Potential competitors include large returns of
pink salmon (@whitehouse2023a, @vulstek2023a), a relatively large
population of Pacific ocean perch (Assessment @hulson2022a,
Appendix 1A), large year classes of juvenile sablefish
(Assessment, @goethel2023a, Appendix 1A).


#### Fishery performance
Trends in effort-weighted fishery CPUE were examined in the ESP
(Appendix 1A) for two seasons, the pre-spawning fishery (A and B
seasons) and the summer/fall fishery (C and D seasons). Fishery
CPUE is either above (A and B seasons) or close to (C and D
seasons) the long-term average, and is very consistent with the
abundance trend of exploitable biomass from the assessment. No
concerns regarding fishery performance were identified and this
element was given a score of 1.


#### Summary and ABC recommendation

```{r}
l1 = "Level 1: No concern"
l2 = "Level 2: Major concern"
l3 = "Level 3: Extreme concern"

data.frame(assess = l2,
           pop = l1,
           enviro = l1,
           fish = l1) %>%
  flextable() %>%
    width(j = ~ assess, width = 1.5) %>%
    width(j = ~ pop, width = 1.5) %>%
    width(j = ~ enviro, width = 1.5) %>%
    width(j = ~ fish, width = 1.5) %>%
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>%
  fontsize(size = 10) %>%
  font(fontname = "Times", part = "all") %>%
  italic(part = "header")
```

Given the overall lack of elevated scores in the risk table, the
author’s recommended ABC is based on the maximum permissible ABC,
resulting in a `r year+1` ABC of `r maxabc` t, which is a 
`r pct.abc.changeF`from the `r year` ABC. The author’s recommended
`r year+2` ABC is `r F(exec_table[12,4])` t. The OFL in 
`r year+1` is `r ofl` t, and the OFL in `r year+2` if the `r year+1`
ABC is taken in `r year+1` is `r F(exec_table[10,4])` t.  

To evaluate the probability that the stock will drop below the
B20% threshold, we projected the stock forward for five years
using the author’s recommended fishing mortality schedule. This
projection incorporates uncertainty in stock status, uncertainty
in the estimate of B20%, and variability in future
recruitment. We then sampled from the probability of future
spawning biomass using Markov chain Monte Carlo (MCMC) using the
no-U-turn sampler available in TMB [@monnahan2018a]. Analysis of
the posterior samples indicates that probability of the stock
dropping below B20% will be negligible through `r year+5`,
conditional upon the model specified here
(Fig. 1.\@ref(fig:depletion)).
